### 2023-10-03 Manuscript Analysis code 0 to 19.Rmd
Updated using ghlt contrast furnction for time interval calculations

---
title: "SEER Medicaid"
author: "Kim Johnson"
date: '2022-06-06'
output: html_document
---

# Open libraries
```{r}
# install.packages("pacman")
pacman::p_load(readr, dplyr, tidyverse, lubridate, survival, survminer, hablar, table1, patchwork, openxlsx, DiagrammeR, lmtest, forestplot, MASS, nnet, broom, forcats, sandwich, mlogit, sjstats, odds.n.ends, labelled, readxl)
```
# Functions used
```{r}
# function to calculate clustered SEs (from: https://stackoverflow.com/questions/73136246/clustered-standard-errors-stars-and-summary-statistics-in-modelsummary-for-mul)
mlogit.clust <- function(model,data,variable) {
  beta <- c(t(coef(model)))
  vcov <- vcov(model)
  k <- length(beta)
  n <- nrow(data)
  max_lev <- length(model$lev)
  xmat <- model.matrix(model)
  # u is deviance residuals times model.matrix
  u <- lapply(2:max_lev, function(x)
    residuals(model, type = "response")[, x] * xmat)
  u <- do.call(cbind, u)
  m <- dim(table(data[,variable]))
  u.clust <- matrix(NA, nrow = m, ncol = k)
  fc <- factor(data[,variable])
  for (i in 1:k) {
    u.clust[, i] <- tapply(u[, i], fc, sum)
  }
  cl.vcov <- vcov %*% ((m / (m - 1)) * t(u.clust) %*% (u.clust)) %*% vcov
  return(cl.vcov = cl.vcov)
}
```

# Import files sent by SEER Medicaid (IMS), get data into right columns, and assign column names
```{r, eval =FALSE}
# # Medicaid enrollment--imports fixed width columns and gives them names
# enroll <- read_fwf('medicaid.enroll.txt', 
#                fwf_cols(patient_id = 15, STATE_CD = 2, YEAR = 4, DOB_MATCH_SCORE_HI = 1, SEX_MATCH_SCORE_HI = 1,
#                        ID_MATCH_SCORE_HI=1, MATCH_SCORE_HI=2, MEDICAID_ELIG_CD1=1, MEDICAID_ELIG_CD2=1,
#                       MEDICAID_ELIG_CD3=1, MEDICAID_ELIG_CD4=1, MEDICAID_ELIG_CD5=1, MEDICAID_ELIG_CD6=1,
#                       MEDICAID_ELIG_CD7=1, MEDICAID_ELIG_CD8=1, MEDICAID_ELIG_CD9=1, MEDICAID_ELIG_CD10=1,
#                         MEDICAID_ELIG_CD11=1, MEDICAID_ELIG_CD12=1), 
#                skip = 0)

# # SEER file (2 is the second dataseet received that includes insurance variables)
# 
# seer<-read_fwf('medicaid.cancer 2.txt', na=".", fwf_positions(
# c(1, 16,18,20,21,22,24,25,28,31,33,35,39,41,45,46,50,51,55,56,57,58,59,62,64,66,67,69, 71, 72,73,74,75,76,77,78,79,80,81,82,83,84,86,88,90,92,94,95,96,97,98,99,100,103,106,109, 111,114,117,120,123,126,129,132,135,138,141,144,147,150,153,156, 159,161,163,165,167,168,174,180,186,188,189,190,192,193,194,195,196,197,199, 200,201,203,205,210,213,216,217,219,224,227,228,229,230,231,233,235,236,238,241,242, 243,248, 253, 254, 255,256, 258,263,264,265,266,267,268,269,270,271,273,279,282,285,288,291,293,295,297,299, 300,301,302,303,304,305,306,308,310,316,347,353,359,365,366,367,368,369,378,380,382,384,388,401,402,403,404,405, 406,408,412,414, 419, 423,425,427,429, 431,433,437,438,439,440,441,442,443,445, 447, 448, 452, 456, 460, 464,465,466,467,468,469,470, 471,472,473,474,475,476,477, 479, 480,485,490,495,500,505, 510,511,512,513,514,515,519,520, 521,524,525,526,530,532),

# c(15,17,19,20,21,23,24,27,30,32,34,38,40,44,45,49,50,54,55,56,57,58,61,63,65,66,68,70,71, 72,73,74,75,76,77,78,79,80,81,82,83,84,86,88,90,93,94,95,96,97,98, 99,102,105,108,110,113,116,119,122,125,128,131,134,137, 140, 143,146,149,152,155,158,160,162,164,166,167,173,179, 185,187,188,189,191, 192,193,194,195,195, 198,199, 200,202,204,209,212,215,216,218,220, 226,227,228,229,230,232,234,235,237,240,241,242,247, 252, 253, 254,255,257, 259, 263,264,265,266,267,268,269,270, 272,273,281,284,287,290,292,294,295,298,299, 300, 301,302,303,304,305,307, 309,311,316,352,358, 364,365,366,367,368,369, 379,381,383,387,391,401,402,403,404,405, 407,411,413,417, 422,424,426,428,430, 432,436,437,438,439,440,441,442,444,446, 447, 451, 455,459,463,464, 465, 466,467,468,469,470, 471,472,473,474,475,476,477,479, 484,489,494,499,504,509,510,511,512,513,514,515,519,520, 523,524,525,529,531,534)))

# colnames(seer)<- c("patient_id", "SEER_registry", "SEERregistrywithCAandGAaswholes", 
# "Louisiana20051stvs2ndhalfofyear", "Marital_
# _at_diagnosis", 
# "Race_ethnicity", "sex", "Agerecodewithsingleages_and_100", "agerecodewithsingle_ages_and_85", 
# "Sequence_number", "Month_of_diagnosis", "Year_of_diagnosis", 
# "Month_of_diagnosis_recode", "Primary_Site", "Laterality", "Histology_ICD_O_2", 
# "Behavior_code_ICD_O_2", "Histologic_Type_ICD_O_3", "Behavior_code_ICD_O_3", 
# "Grade", "Diagnostic_Confirmation", "Type_of_Reporting_Source", 
# "EOD_10_size_1988_2003", "EOD_10_extent_1988_2003", "EOD10Prostatepath_ext_1995_2003", 
# "EOD_10_nodes_1988_2003", "Regional_nodes_positive_1988", "Regional_nodes_examined_1988", 
# "Expanded_EOD_1_CP53_1973_1982", "Expanded_EOD_2_CP54_1973_1982", 
# "Expanded_EOD_3_CP55_1973_1982", "Expanded_EOD_4_CP56_1973_1982", 
# "Expanded_EOD_5_CP57_1973_1982", "Expanded_EOD_6_CP58_1973_1982", 
# "Expanded_EOD_7_CP59_1973_1982", "Expanded_EOD_8_CP60_1973_1982",
# "Expanded_EOD_9_CP61_1973_1982", "Expanded_EOD_10_CP62_1973_1982",
# "Expanded_EOD_11_CP63_1973_1982", "Expanded_EOD_12_CP64_1973_1982",
# "Expanded_EOD_13_CP65_1973_1982", "V2Digit_NS_EOD_part_1_1973_1982",
# "V2Digit_NS_EOD_part_2_1973_1982", "V2Digit_SS_EOD_part_1_1973_1982",
# "V2Digit_SS_EOD_part_2_1973_1982", "EOD_4_size_1983_1987", "EOD_4_extent_1983_1987",
# "EOD_4_nodes_1983_1987", "Coding_system_EOD_1973_2003", "Tumor_marker_1_1990_2003",
# "Tumor_marker_2_1990_2003", "Tumor_marker_3_1998_2003", "CS_tumor_size_2004_2015",
# "CS_extension_2004_2015", "CS_lymph_nodes_2004_2015", "CS_mets_at_dx_2004_2015",
# "CSsitespecificfactor12004varyin", "CSsitespecificfactor22004varyin",
# "CSsitespecificfactor32004varyin", "CSsitespecificfactor42004varyin",
# "CSsitespecificfactor52004varyin", "CSsitespecificfactor62004varyin",
# "CSsitespecificfactor72004varyin", "CSsitespecificfactor82004varyin",
# "CSsitespecificfactor92004varyin", "CSsitespecificfactor102004varyi",
# "CSsitespecificfactor112004varyi", "CSsitespecificfactor122004varyi",
# "CSsitespecificfactor132004varyi", "CSsitespecificfactor152004varyi",
# "CSsitespecificfactor162004varyi", "CSsitespecificfactor252004varyi",
# "Derived_AJCC_T_6th_ed_2004_2015", "Derived_AJCC_N_6th_ed_2004_2015",
# "Derived_AJCC_M_6th_ed_2004_2015", "DerivedAJCCStageGroup6thed20042",
# "SEERCombinedSummaryStage2000200", "CSversioninputoriginal2004_2015",
# "CS_version_derived_2004_2015", "CSversioninputcurrent_2004_2015",
# "RX_Summ_Surg_Prim_Site_1998", "RX_Summ_Scope_Reg_LN_Sur_2003",
# "RX_Summ_Surg_Oth_Reg_Dis_2003", "RXSummReg_LN_Examined_1998_2002",
# "RX_Summ_Systemic_Surg_Seq", "RX_Summ_Surg_Rad_Seq", "Reasonnocancer_directed_surgery",
# "Radiation_recode", "Chemotherapy_recode_yes_no_unk", "Sitespecificsurgery19731997vary",
# "Scopeofreglymphndsurg_1998_2002", "Surgeryofothregdissites19982002",
# "Record_number_recode", "Age_recode_with_1_year_olds", "Site_recode_ICD_O_3_WHO_2008",
# "ICCCsiterecode_ICD_O_3_WHO_2008", "ICCCsiterecextendedICDO3WHO2008",
# "Behavior_recode_for_analysis", "Histologyrecode_broad_groupings",
# "Histologyrecode_Brain_groupings", "TNM_7_CS_v0204_Schema", "Race_recode_White_Black_Other",
# "Race_recode_W_B_AI_API", "OriginrecodeNHIAHispanicNonHisp",
# "SEER_historic_stage_A_1973_2015", "AJCCstage_3rd_edition_1988_2003",
# "SEERmodifiedAJCCstage3rd1988200", "Firstmalignantprimary_indicator",
# "state", "county", "PRCDA_2017", "PRCDA_Region", "COD_to_site_recode",
# "COD_to_site_rec_KM", "Vital
# recodestudycutoffuse", "IHS_Link",
# "Summary_stage_2000_1998", "AYA_site_recode_WHO_2008", "Lymphomasubtype_recode_WHO_2008",
# "SEERcausespecificdeathclassific", "SEERothercauseofdeathclassifica",
# "CSTumor_Size_Ext_Eval_2004_2015", "CS_Reg_Node_Eval_2004_2015",
# "CS_Mets_Eval_2004_2015", "Primary_by_international_rules", "ER
# RecodeBreastCancer_1990",
# "PR
# RecodeBreastCancer_1990", "CS_Schema_AJCC_6th_Edition",
# "LymphvascularInvasion2004varyin", "Derived_AJCC_T_7th_ed_2010",
# "Derived_AJCC_N_7th_ed_2010", "Derived_AJCC_M_7th_ed_2010", "DerivedAJCCStageGroup7thed_2010",
# "BreastAdjustedAJCC6thT1988_2015", "BreastAdjustedAJCC6thN1988_2015",
# "BreastAdjustedAJCC6thM1988_2015", "BreastAdjustedAJCC6thStage19882",
# "Derived_HER2_Recode_2010", "Breast_Subtype_2010", "LymphomaAnnArborStage_1983_2015",
# "SEERCombinedMetsat_DX_bone_2010", "SEERCombinedMetsatDX_brain_2010",
# "SEERCombinedMetsatDX_liver_2010", "SEERCombinedMetsat_DX_lung_2010",
# "TvaluebasedonAJCC_3rd_1988_2003", "NvaluebasedonAJCC_3rd_1988_2003",
# "MvaluebasedonAJCC_3rd_1988_2003", "RadiationtoBrainorCNSRecode1988",
# "Census_Tract_1990", "Census_Tract_2000", "Census_Tract_2010",
# "Census_Coding_System", "Census_Tract_Certainty_1990", "Census_Tract_Certainty_2000",
# "Census_Tract_Certainty_2010", "Census_Tract_Poverty_Indicator",
# "Rural_Urban_Continuum_Code_1993", "Rural_Urban_Continuum_Code_2003",
# "Rural_Urban_Continuum_Code_2013", "Health_Service_Area", "HealthService_Area_NCI_Modified",
# "Derived_SS1977_flag", "Derived_SS2000_flag", "Firstcourseofreconstruct1998200",
# "Radiation", "RadiationtoBrainorCNS_1988_1997", "SEER_DateofDeath_Month",
# "SEER_DateofDeath_Year", "Month_of_last_follow_up_recode", "Year_of_last_follow_up_recode",
# "Year_of_birth", "Date_of_diagnosis_flag", "Date_therapy_started_flag",
# "Date_of_birth_flag", "Date_of_last_follow_up_flag", "Month_therapy_started",
# "Year_therapy_started", "Other_cancer_directed_therapy", "Derived_AJCC_flag",
# "Derived_SS1977", "Derived_SS2000", "SEER_Summary_stage_1977_9500",
# "SEER_Summary_stage_2000_0103", "NAPIIA_Derived_API_Race", "Primary_Payer_at_DX",
# "COD_ICD_code", "Cause_of_Death_ICD_8_or_9", "Cause_of_Death_ICD_10",
# "Recode_ICD_0_2_to_9", "Recode_ICD_0_2_to_10", "NHIA_Derived_Hisp_Origin",
# "Age_site_edit_override", "Sequencenumber_dx_conf_override",
# "Site_type_lat_seq_override", "Surgerydiagnostic_conf_override",
# "Site_type_edit_override", "Histology_edit_override", "Report_source_sequence_override",
# "Seq_ill_defined_site_override", "LeukLymphdxconfirmationoverride",
# "Site_behavior_override", "Site_EOD_dx_date_override", "Site_laterality_EOD_override",
# "Site_laterality_morph_override", "Insurance_Recode_2007", "Yost_ACS_2006_2010",
# "Yost_ACS_2010_2014", "Yost_ACS_2013_2017", "Yost_ACS_2006_2010_State_based",
# "Yost_ACS_2010_2014_State_based", "Yost_ACS_2013_2017_State_based",
# "Yost_ACS_2006_2010_quintile", "Yost_ACS_2010_2014_quintile",
# "Yost_ACS_2013_2017_quintile", "YostACS20062010quintileStatebas",
# "YostACS20102014quintileStatebas", "YostACS20132017quintileStatebas",
# "HPV_recode_2010", "In_Medicaid_Flag", "OncotypeDXBreastRecurrenceScore",
# "OncotypeDXRSgroupRS18RS1830RS30", "OncotypeDXreasonno_score_linked",
# "Oncotype_DX_year_of_test_report", "OncotypeDX_month_of_test_report",
# "OncotypeDXmonthssince_diagnosis")

# hit control shift c while highlighting above to uncomment/comment section
```

# Create dataframes for Medicaid enrollment and SEER for export that are limited to those diagnosed at 0 to 39 years 
# Exclude those over 39 years old and export data to make the datasets more manageeable on import
```{r}
# exclude those 40 and over from seer cancer filefirst
# seer_0to39<-seer %>%
#   filter(Agerecodewithsingleages_and_100<40)

# get ids from seer file to limit enrollment file to those ids

# ids<-seer_0to39$patient_id

# enroll_seerids<-enroll %>% # limit Medicaid enrollment file to SEER ids
 # filter(patient_id %in% ids)
```

# Export these data as csv files for quicker import
```{r}
# write.csv(seer_0to39,"seer0to39.csv", row.names = FALSE)

# write.csv(enroll_seerids,"enroll_seerids.csv", row.names = FALSE)
```

# Import files created with text commented out above for analysis
```{r}
seer0to39<-read_csv("/Volumes/kijohnson/Active/seer0to39.csv", show_col_types = FALSE) # this is 0 to 39 cancer patient data
enroll_seerids<-read_csv("/Volumes/kijohnson/Active/enroll_seerids.csv", show_col_types = FALSE) # this is 0 to 39 enrollee data
mcaid<- read_xlsx("/Volumes/kijohnson/Active/mcaid0to39KJ.xlsx") # created in continuity coding file
```

# Identify ids to exclude where patient has Medicaid files in multiple states (this is used below)
```{r}
df <- enroll_seerids %>% 
   find_duplicates(patient_id, YEAR) # this identifies patients in multiple states in the same year

x <- unique(df$patient_id) # 841 vector of unique ids who had enrollment in multiple states to exclude in exclusion section
```

# Reduce cancer dataset to variables needed for analysis
```{r}
# select variables
seer0to39<-seer0to39 %>%
  dplyr::select("patient_id", "SEER_registry",  "Race_ethnicity", "sex", "Agerecodewithsingleages_and_100",
"Sequence_number", "Month_of_diagnosis", "Year_of_diagnosis",
"Month_of_diagnosis_recode", "Primary_Site", "Laterality", "Histology_ICD_O_2",
"Behavior_code_ICD_O_2", "Histologic_Type_ICD_O_3", "Behavior_code_ICD_O_3",
"Grade", "Diagnostic_Confirmation", "Type_of_Reporting_Source", 
"Derived_AJCC_T_6th_ed_2004_2015", 
"Derived_AJCC_N_6th_ed_2004_2015",
"Derived_AJCC_M_6th_ed_2004_2015", 
"DerivedAJCCStageGroup6thed20042",
"SEERCombinedSummaryStage2000200", 
"Age_recode_with_1_year_olds", "Site_recode_ICD_O_3_WHO_2008",
"ICCCsiterecode_ICD_O_3_WHO_2008", "ICCCsiterecextendedICDO3WHO2008",
"Behavior_recode_for_analysis", "Histologyrecode_broad_groupings", "Race_recode_White_Black_Other",
"Race_recode_W_B_AI_API", "OriginrecodeNHIAHispanicNonHisp",
"SEER_historic_stage_A_1973_2015", "AJCCstage_3rd_edition_1988_2003",
"SEERmodifiedAJCCstage3rd1988200", "Firstmalignantprimary_indicator",
"state", "county", "PRCDA_2017", "PRCDA_Region", "COD_to_site_recode",
"COD_to_site_rec_KM", "Vitalstatusrecodestudycutoffuse", "IHS_Link",
"Summary_stage_2000_1998", "AYA_site_recode_WHO_2008", 
"SEERcausespecificdeathclassific", "SEERothercauseofdeathclassifica",
"Derived_AJCC_T_7th_ed_2010", "Derived_AJCC_N_7th_ed_2010", "Derived_AJCC_M_7th_ed_2010", "DerivedAJCCStageGroup7thed_2010", "Census_Tract_1990", "Census_Tract_2000", "Census_Tract_2010",
"Census_Coding_System", "Census_Tract_Certainty_1990", "Census_Tract_Certainty_2000",
"Census_Tract_Certainty_2010", "Census_Tract_Poverty_Indicator",
"Rural_Urban_Continuum_Code_1993", "Rural_Urban_Continuum_Code_2003",
"Rural_Urban_Continuum_Code_2013", "Health_Service_Area", "HealthService_Area_NCI_Modified",
"SEER_DateofDeath_Month", "SEER_DateofDeath_Year", "Month_of_last_follow_up_recode", "Year_of_last_follow_up_recode", "Year_of_birth", "Date_of_diagnosis_flag", "Date_therapy_started_flag",
"Date_of_birth_flag", "Date_of_last_follow_up_flag", "Month_therapy_started",
"Year_therapy_started",  "Derived_AJCC_flag",
"SEER_Summary_stage_1977_9500",
"SEER_Summary_stage_2000_0103", "NAPIIA_Derived_API_Race", "Primary_Payer_at_DX",
"COD_ICD_code", "Cause_of_Death_ICD_8_or_9", "Cause_of_Death_ICD_10",
"Recode_ICD_0_2_to_9", "Recode_ICD_0_2_to_10", "NHIA_Derived_Hisp_Origin",
"Insurance_Recode_2007", "Yost_ACS_2006_2010", "Yost_ACS_2010_2014", "Yost_ACS_2013_2017", "Yost_ACS_2006_2010_State_based", "Yost_ACS_2010_2014_State_based", "Yost_ACS_2013_2017_State_based",
"Yost_ACS_2006_2010_quintile", "Yost_ACS_2010_2014_quintile",
"Yost_ACS_2013_2017_quintile", "YostACS20062010quintileStatebas",
"YostACS20102014quintileStatebas", "YostACS20132017quintileStatebas",
"HPV_recode_2010", "In_Medicaid_Flag")
```

# Data management of variables 
```{r}
# recode demographic variables
# Age, race, hispanic, sex  cause of death, cause specfic death
seer0to39<-seer0to39 %>%
  mutate(age_cat=case_when(Age_recode_with_1_year_olds %in% c(0,1) ~ 0, # 0 to 4
                                Age_recode_with_1_year_olds == 2 ~1, # 5 to 9
                                Age_recode_with_1_year_olds == 3 ~2, # 10 to 14
                                Age_recode_with_1_year_olds == 4 ~3, # 15 to 19
                                Age_recode_with_1_year_olds %in% c(5,6,7,8) ~ 4)) %>% # 20 to 39

         mutate(race_cat=case_when(Race_recode_W_B_AI_API == 1 ~0, # White
                                Race_recode_W_B_AI_API == 2 ~1, # Black
                                Race_recode_W_B_AI_API == 3 ~2, # AIAN
                                Race_recode_W_B_AI_API == 4 ~3, # API
                                Race_recode_W_B_AI_API == 9 ~4)) %>% # Unknown

        mutate(hispanic = case_when(OriginrecodeNHIAHispanicNonHisp == 0 ~0, # Non-Hispanic
                                    OriginrecodeNHIAHispanicNonHisp == 1 ~1)) %>% # Hispanic

        mutate(race_ethnicity = case_when(race_cat == 0 & hispanic == 0 ~ 0, # Non-Hispanic White
                                race_cat == 1 & hispanic == 0 ~ 1, # Non-Hispanic Black
                                race_cat == 3 & hispanic == 0 ~ 2, # Non-Hispanic API
                                race_cat %in% c(2,4) & hispanic == 0 ~ 3, # Non-Hispanic Other
                                race_cat %in% c(0:4) & hispanic == 1 ~ 4)) %>% # Hispanic

        mutate(race_cat2 = case_when(race_cat == 0 ~ 0, # Non-Hispanic White
                                     race_cat == 1 ~ 1, # Non-Hispanic Black
                                     race_cat %in% c(2,3,4,5) ~ 2)) %>% # Non-Hispanic Other

        mutate(sex = case_when(sex == 1 ~ 0, # "Male"
                                sex == 2 ~ 1, # "Female"
                                sex == 9 ~ 2)) %>% # "Unknown"

        mutate(can_dead = case_when(SEERcausespecificdeathclassific == 0 ~0, # Alive or dead of other cause
                                SEERcausespecificdeathclassific == 1 ~1)) # Dead from cancer
# note 8 an 9 are missing or N/A

table(seer0to39$SEERcausespecificdeathclassific, useNA = "always")

# Factor variables
seer0to39$age_cat<-factor(seer0to39$age_cat, levels=c(0:4), labels=c("0-4", "5-9", "10-14", "15-19", "20-39"))
seer0to39$race_cat<-factor(seer0to39$race_cat, levels=c(0:4), labels=c("White", "Black", "AIAN", "API", "Unknown"))
seer0to39$hispanic<-factor(seer0to39$hispanic, levels=c(0:1), labels=c("Non-Hispanic", "Hispanic"))
seer0to39$race_ethnicity<-factor(seer0to39$race_ethnicity, levels=c(0:4), labels=c("Non-Hispanic White", "Non-Hispanic Black", "Non-Hispanic API", "Non-Hispanic Other",  "Hispanic")) 
seer0to39$sex <- factor(seer0to39$sex, levels = c(0:2), labels = c("Male", "Female", "Not Stated"))

# check variables
table(seer0to39$age_cat, useNA = "always")
table(seer0to39$race_cat, useNA = "always")
table(seer0to39$hispanic, useNA = "always")
table(seer0to39$race_ethnicity, useNA = "always")
table(seer0to39$can_dead, useNA = "always")
table(seer0to39$sex, useNA = "always")

# calculate follow-up time from month of dx, recode year of dx TO Month_of_last_follow_up_recode, Year_of_last_follow_up_recode. The difference between these variable dates would provide months of follow-up
summary(seer0to39$Month_of_diagnosis_recode)
summary(seer0to39$Year_of_diagnosis)
summary(as.numeric(seer0to39$Month_of_last_follow_up_recode))
seer0to39$Month_of_last_follow_up_recode<-as.numeric(seer0to39$Month_of_last_follow_up_recode) # make variable numeric
seer0to39$SEER_DateofDeath_Month<-as.numeric(seer0to39$SEER_DateofDeath_Month) # make variable numeric
summary(seer0to39$Month_of_last_follow_up_recode)
summary(seer0to39$Year_of_last_follow_up_recode)
summary(seer0to39$SEER_DateofDeath_Month)
summary(seer0to39$SEER_DateofDeath_Year)

# assign a date of diagnosis using 15 as the day because there is no day of dx so assume mean is 15
seer0to39$proxy_day_of_dx <-15
seer0to39$proxy_day_of_fu <-15
seer0to39$proxy_day_of_dth <-15

seer0to39<-seer0to39 %>%
  mutate(proxy_dx_date = make_date(Year_of_diagnosis, Month_of_diagnosis_recode, proxy_day_of_dx)) %>% # make date of dx
  mutate(proxy_fu_date = make_date(Year_of_last_follow_up_recode, Month_of_last_follow_up_recode, proxy_day_of_fu)) %>% # make date of follow-up
  mutate(proxy_dth_date = make_date(SEER_DateofDeath_Year, SEER_DateofDeath_Month, proxy_day_of_dth)) # make date of death

# calculate the difference in days between the estimated date of dx and estimated date of last follow-up for non-cases

# non-cases
seer0to39_noncases <- seer0to39 %>%
  filter(can_dead == 0|is.na(can_dead)) # used last follow-up for missing can_dead
seer0to39_noncases$survday<-as.numeric(difftime(seer0to39_noncases$proxy_fu_date, seer0to39_noncases$proxy_dx_date, units=c("days")))

# cases
seer0to39_cases <- seer0to39 %>%
  filter(can_dead == 1)
seer0to39_cases$survday <-as.numeric(difftime(seer0to39_cases$proxy_dth_date, seer0to39_cases$proxy_dx_date, units=c("days")))

summary(seer0to39_noncases$proxy_fu_date)
summary(seer0to39_cases$proxy_fu_date)

# bind these back together
seer0to39 <- rbind(seer0to39_cases, seer0to39_noncases)

# estimate follow-up months by dividing by the average number of days in a month
seer0to39$fumo<-seer0to39$survday/30.437

summary(seer0to39$fumo)

# cancer type coding
table(seer0to39$ICCCsiterecode_ICD_O_3_WHO_2008)

# ICCC3 (https://seer.cancer.gov/iccc/iccc-who2008.html)
seer0to39<-seer0to39 %>%
  mutate(can_type = case_when(
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(11, 12,13,14,15)~0,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(21,22,23,24,25)~1,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(31,32,33,34,35,36)~2,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(41,42) ~3,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(50) ~4,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(61,62,63) ~5,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(71,72,73) ~6,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(81,82, 83,84,85) ~7,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(91,92,93,94,95)~8,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(101,102,103, 104,105) ~9,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(111,112,113,114,115,116) ~10,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(121,122)~11))

seer0to39$can_type<-factor(seer0to39$can_type, levels=c(0:11), labels=c(
  "I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases",
  "II. Lymphomas and reticuloendothelial neoplasms",
  "III. CNS and miscellaneous intracranial and intraspinal neoplasms",
  "IV. Neuroblastoma and other peripheral nervous cell tumors",
  "V. Retinoblastoma",
  "VI. Renal tumors",
  "VII. Hepatic tumors",
  "VIII. Malignant bone tumors",
  "IX. Soft tissue and other extraosseous sarcomas",
  "X. Germ cell tumors, trophoblastic tumors, and neoplasms of gonads",
  "XI. Other malignant epithelial neoplasms and malignant melanomas",
  "XII. Other and unspecified malignant neoplasms"))

table(seer0to39$can_type, seer0to39$ICCCsiterecode_ICD_O_3_WHO_2008, useNA= "always")

table(seer0to39$AYA_site_recode_WHO_2008)

# use Yost index as a measure of poverty
table(seer0to39$Yost_ACS_2006_2010_quintile,  useNA = "always")

class(seer0to39$Yost_ACS_2006_2010_quintile)

seer0to39$Yost_ACS_2006_2010_quintile <-factor(seer0to39$Yost_ACS_2006_2010_quintile, levels = c(1:5), labels = c(
        "Group 1 (lowest SES)",
        "Group 2",
        "Group 3",
        "Group 4",
        "Group 5 (highest SES)"))

# classify stage
table(seer0to39$SEERCombinedSummaryStage2000200, useNA = "always")
seer0to39 <- seer0to39 %>%
  mutate(stage = case_when(SEERCombinedSummaryStage2000200 %in% c(0,1) ~ 0,
                           SEERCombinedSummaryStage2000200 %in% c(2,3,4,5) ~ 1,
                           SEERCombinedSummaryStage2000200 %in% c(7) ~ 2))

seer0to39$stage <- factor(seer0to39$stage, levels = c(0:2), labels = c("In situ/localized", "Regional", "Distant"))

table(seer0to39$stage, useNA ="always")

table(seer0to39$SEER_DateofDeath_Year, useNA = "always")

table(seer0to39$can_type)

```

# test code
```{r}

# create categories for AYA codes
class(seer0to39$AYA_site_recode_WHO_2008) # numeric

seer0to39 <- seer0to39 %>%
  mutate(can_typeAYA = case_when(AYA_site_recode_WHO_2008 %in% c(1,2,3,4) ~ 0,
                                 AYA_site_recode_WHO_2008 %in% c(5,6) ~ 1,
                                 AYA_site_recode_WHO_2008 %in% c(7,8,9,10,11,12,13,14,15,16) ~ 2,
                                 AYA_site_recode_WHO_2008 %in% c(17,18,19,20) ~ 3,
                                 AYA_site_recode_WHO_2008 %in% c(21,22,23,24,25) ~ 4,
                                 AYA_site_recode_WHO_2008 %in% c(26,27,28) ~ 5,
                                 AYA_site_recode_WHO_2008 %in% c(29,30) ~ 6,
                                 AYA_site_recode_WHO_2008 %in% c(31,32,33,34,35,36,37,38,39,40,41,42,43,44,45,46,47,48) ~ 7,
                                 AYA_site_recode_WHO_2008 %in% c(49,50,51,52,53,54,55) ~ 8,
                                 AYA_site_recode_WHO_2008 %in% c(55,56) ~ 9,
                                 AYA_site_recode_WHO_2008 == 99 ~ 10))

seer0to39$can_typeAYA <- factor(seer0to39$can_typeAYA, levels = c(0:10), labels = c(
  "1. Leukemias",
  "2. Lymphomas",
  "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)",
  "4. Osseous & Chondromatous Neoplasms",
  "5. Soft Tissue Sarcomas",
  "6. Germ Cell and Trophoblastic Neoplasms",
  "7. Melanoma and Skin Carcinomas",
  "8. Carcinomas",
  "9. Miscellaneous specified neoplasms, NOS",
  "10. Unspecified Malignant Neoplasms",
  "Unclassified"))
table(seer0to39$can_typeAYA, useNA = "always")

# categorize brain tumors according to AYA definition

seer0to39 <- seer0to39 %>%
  mutate(brain = case_when (can_typeAYA == "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)"  & AYA_site_recode_WHO_2008 %in% c(7,8,9) ~ 0, # 3.1 Astrocytoma
                            can_typeAYA == "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)" & AYA_site_recode_WHO_2008 %in% c(10) ~ 1, # 3.2 Other glioma
                            can_typeAYA == "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)" & AYA_site_recode_WHO_2008 %in% c(11) ~ 2, # 3.3 Ependymoma
                            can_typeAYA == "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)" & AYA_site_recode_WHO_2008 %in% c(12,13) ~ 3, # 3.4 Medulloblastoma and other PNET
                            can_typeAYA == "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)" & AYA_site_recode_WHO_2008 %in% c(14) ~ 4, # 3.5 Other specified intracranial and instraspinal neoplasms
                            can_typeAYA == "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)" & AYA_site_recode_WHO_2008 %in% c(15,16) ~ 5)) # 3.6 Unspecified intracranial and intraspinal neoplasms
seer0to39$brain
table(seer0to39$brain, useNA = "always")

seer0to39$brain <- factor(seer0to39$brain, levels = c(0:5), labels = c(
  "3.1 Astrocytoma",
  "3.2 Other glioma",
  "3.3 Ependymoma",
  "3.4 Medulloblastoma and other PNET",
  "3.5 Other specified intracranial and instraspinal neoplasms",
  "3.6 Unspecified intracranial and intraspinal neoplasms"))
table(seer0to39$brain, useNA = "always")

df <- seer0to39 %>%
  filter(can_type == "III. CNS and miscellaneous intracranial and intraspinal neoplasms"|can_typeAYA == "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)") %>%
  group_by(can_type, can_typeAYA)%>%
  tally()
```

# Exclusions

# make separate dataframe for 0 to 19 
```{r}
seer0to19 <- seer0to39 %>%
  filter(age_cat %in% c("0-4", "5-9", "10-14", "15-19")) # 34,196

seer0to19$age_cat <- droplevels(seer0to19$age_cat)
```

# limit to first cancers
```{r}
seer0to19 <- seer0to19 %>%
  filter(Firstmalignantprimary_indicator == 1) # 32,396
```
# limit to sequence number 01 or 00
```{r}
seer0to19 <- seer0to19 %>%
  filter(Sequence_number %in% c("00", "01")) # 32,380
```

# Exclude those diagnosed before 2006 (a handful)
```{r}
seer0to19 <- seer0to19 %>%
  filter(Year_of_diagnosis >=2006) # 32,380
```

# Exclude multiple state Medicaid ids within a single year defined above
```{r}
seer0to19 <- seer0to19 %>% 
  filter(!patient_id %in% x) # 31,562
```

# Exclude those with Medicaid cancer registry match scores below  5
```{r}
match <- enroll_seerids %>%
  filter(MATCH_SCORE_HI <5)

# create list of unique ids to exclude
ids_2_exclude <- unique(match$patient_id)

seer0to19 <- seer0to19 %>%
  filter(!patient_id %in% ids_2_exclude) # 31,410
```

# Exclude those with missing data on variables used in the analysis
```{r}
seer0to19<-seer0to19 %>%
  filter(!is.na(fumo)) #31,410

seer0to19<-seer0to19 %>%
  filter(!is.na(can_dead)) # 31,247

seer0to19<-seer0to19 %>%
  filter(!is.na(In_Medicaid_Flag)) # 31,247

seer0to19<-seer0to19 %>%
  filter(!is.na(race_ethnicity)) # 31,247

seer0to19<-seer0to19 %>%
  filter(!is.na(age_cat)) # 31,247

seer0to19<-seer0to19 %>%
  filter(!is.na(Yost_ACS_2006_2010_quintile)) #30,654

seer0to19b <- seer0to19

missing_data_percent  <- 100- (30654/31410)*100
missing_data_percent  # <2.5%
```

# merge in mcont (medicaid enrollment data and recode 0 = Not linked, 1 = continuous, 2 = discontinuous, 3 = Other)
```{r}
seer0to19 <- left_join(seer0to19b, mcaid, by = "patient_id")

table(seer0to19$mcontb, useNA = "always")
table(mcaid$mcontb, useNA = "always")
table(seer0to19$In_Medicaid_Flag)

# recode a category for not in medicaid
seer0to19 <- seer0to19 %>%
  mutate(mcont66 = case_when(In_Medicaid_Flag == 0 ~ 0, # not linked
                             mcontb == 0 ~ 1, # continuous
                             mcontb == 1 ~ 2, # discontinuous
                             mcontb == 2 ~ 3)) # Other
table(seer0to19$mcont66, useNA = "always")

table(seer0to19$mcont66, seer0to19$In_Medicaid_Flag, useNA = "always")

# create factor variable for analyses
seer0to19$mcont66 <- factor(seer0to19$mcont66, levels = c(0:3), labels = c("Not enrolled", "Continuous", "Discontinuous", "Other"))
```

# check some variables
```{r}
table(seer0to19$can_type, seer0to19$SEERCombinedSummaryStage2000200, useNA = "always")

table(seer0to19$ICCCsiterecode_ICD_O_3_WHO_2008, seer0to19$can_type,  useNA= "always")
```

# FIGURE 1
```{r}
Figure1<-grViz("digraph flowchart {
      # node definitions with substituted label text
      node [fontname = Helvetica, shape = rectangle, fontsize=16]        
      tab1 [label = '@@1']
      tab2 [label = '@@2']
      tab3 [label = '@@3']
      tab4 [label = '@@4']
      tab5 [label = '@@5']
      tab6 [label = '@@6']
      tab7 [label = '@@7']
      # edge definitions with the node IDs
      tab1 -> tab2 -> tab3 -> tab4 -> tab5 -> tab6 -> tab7;
      }
      [1]: 'Records received for patients diagnosed with malignant cancers at ages 0 to 19 years from 2006-2013 n = 34,196'
      [2]: 'Excluding 1800 individuals whose cancer was not their first malignant primary was n = 32,396' 
      [3]: 'Excluding 16 individuals whose cancer sequence number was not in 0 or 1 n = 32,380'
      [4]: 'Excluding 818 individuals linked to multiple state Medicaid files n = 31,562'
      [5]: 'Excluding 152 individuals with Medicaid enrollment match scores <5 n = 31,410'
      [6]: 'Excluding 163 individuals with missing data on cancer specific cause of death n = 31,247'
      [7]: 'Excluding 593 individuals individuals with missing data on Yost ACS 2006 to 2010 Quintile n = 30,654'
      ")
Figure1
```

# TABLE 1
```{r}
# 0 to 19
label(seer0to19$age_cat)<-"Age category"
label(seer0to19$sex)<-"Sex"
label(seer0to19$race_ethnicity)<-"Race/ethnicity category"
label(seer0to19$fumo)<-"Follow-up months"
label(seer0to19$can_type)<-"Cancer type"
label(seer0to19$Yost_ACS_2006_2010_quintile) <- "Yost Index 2006 to 2010"
label(seer0to19$stage) <- "Stage"
label(seer0to19$mcont66) <- "Medicaid enrollment continuity"

# make can_dead a factor variable just for this table
seer0to19$can_dead2<-factor(seer0to19$can_dead, levels=c(0,1), labels= c("Alive", "Deceased"))
table1(~as.factor(seer0to19$SEERcausespecificdeathclassific)|mcont66, seer0to19)

label(seer0to19$can_dead2) <- "Cancer vital status"
seer0to19$In_Medicaid_Flag<-factor(seer0to19$In_Medicaid_Flag, levels=c(0,1), labels= c("Not Enrolled", "Enrolled"))
label(seer0to19$In_Medicaid_Flag) <- "Enrolled in Medicaid"

table(seer0to19$can_dead2)
table(seer0to19$In_Medicaid_Flag)

table1(~age_cat + sex  + race_ethnicity + Yost_ACS_2006_2010_quintile + In_Medicaid_Flag  + stage + fumo  + can_type + can_dead2|mcont66, droplevels = TRUE, seer0to19)

table1(~age_cat + sex  + race_ethnicity + Yost_ACS_2006_2010_quintile + In_Medicaid_Flag  + stage + fumo  + can_type + can_dead2|mcont66, droplevels = TRUE, seer0to19 %>% filter(Agerecodewithsingleages_and_100 !=18 ))

table1(~mcont66|age_cat, droplevels = TRUE, seer0to19%>% filter(Agerecodewithsingleages_and_100 <18 ))
```

# TABLE 2: Stage analysis (unadjusted)
```{r}
# exclude leukemia cases from stage analysis and those with missing stage
seer0to19_stage <- seer0to19 %>% 
  filter(can_type != "I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases") # this also excludes those with missing cancer type to yield 22,386

# the mlogit.cluster function does not like labels
seer0to19_stage <- seer0to19_stage %>% 
  remove_labels()

# exclude those with missing stage and redefine some variables so the function doesn't throw an error.
seer0to19_stage <- seer0to19_stage %>%
  filter(!is.na(stage)) %>% # this yields 21,502
  select(stage, In_Medicaid_Flag, SEER_registry, mcont66, race_cat2, age_cat, Yost_ACS_2006_2010_quintile, state, can_type) %>%
  mutate(stage2 = case_when(stage == "In situ/localized" ~ "0: In situ/localized",
                            stage == "Regional" ~ "1: Regional",
                            stage == "Distant" ~  "2: Distant"))

# function also likes it to only be a data.frame and nothing else
seer0to19_stage <- as.data.frame(seer0to19_stage)

# In_Medicaid_Flag unadjusted model
model1 <- multinom(stage2 ~ In_Medicaid_Flag, data=seer0to19_stage, trace = FALSE)
model1
summary(model1)

# extract coefficients with non-clustered SEs from the model and exponentiate for comparison 
exp(coef(model1))

# CI(95%)
exp(confint(model1, level=0.95))

# get coefficients, variance, clustered standard errors, and p values
b <- c(t(coef(model1)))
var <- mlogit.clust(model1, seer0to19_stage, "state") # get variance with function
se <- sqrt(diag(var))
p <- (1-pnorm(abs(b/se))) * 2

lowCI <- exp(b - 1.96*se)
highCI <-exp(b + 1.96*se)
df <- as.data.frame(exp(b))
df2<-as.data.frame(lowCI)
df3<-as.data.frame(highCI)
df4 <- cbind(df, df2, df3) # this is the result dataset

# relabel columns and dropt intercept rows
names(df4)[1:3]<-c("OR", "lowCI", "highCI")
df4

df4<-df4[-c(1,3),]
df4 <- tibble::rownames_to_column(df4, "Category") # Apply rownames_to_column

# Replace multiple strings at a time
rep_str = c('1: Regional:In_Medicaid_FlagEnrolled'='Regional','2: Distant:In_Medicaid_FlagEnrolled'='Distant')
rep_str2 = c('1: Regional:In_Medicaid_FlagEnrolled'='Enrolled','2: Distant:In_Medicaid_FlagEnrolled'='Enrolled')
df4$Stage<- str_replace_all(df4$Category, rep_str)
df4$Category<- str_replace_all(df4$Category, rep_str2) # this is the dataframe for summary results

# mcont66 model
model2 <- multinom(stage2 ~ mcont66, data=seer0to19_stage)
summary(model2)

# extract coefficients from the model and exponentiate
exp(coef(model2))

# CI(95%)
exp(confint(model2, level=0.95))

b <- c(t(coef(model2)))
var <- mlogit.clust(model2, seer0to19_stage, "state")
se <- sqrt(diag(var))

lowCI <- exp(b - 1.96*se)
highCI <-exp(b + 1.96*se)

df5 <- as.data.frame(exp(b))
df6<-as.data.frame(lowCI)
df7<-as.data.frame(highCI)
df8 <- cbind(df5, df6, df7) # this is the result dataset
# relabel columns and drop intercept rows

names(df8)[1:3]<-c("OR", "lowCI", "highCI")
df8

df8<-df8[-c(1,5),]
df8 <- tibble::rownames_to_column(df8, "Category") # Apply rownames_to_column
df8

# Replace multiple strings at a time
rep_str = c('1: Regional:mcont66Continuous'='Regional','1: Regional:mcont66Discontinuous'='Regional','1: Regional:mcont66Other'='Regional', '2: Distant:mcont66Continuous'='Distant','2: Distant:mcont66Discontinuous'='Distant','2: Distant:mcont66Other'='Distant')
df8$Stage<- str_replace_all(df8$Category, rep_str)
df8
rep_str2 = c('1: Regional:mcont66Continuous'='Continuous','1: Regional:mcont66Discontinuous'='Discontinuous','1: Regional:mcont66Other'='Other', '2: Distant:mcont66Continuous'='Continuous','2: Distant:mcont66Discontinuous'='Discontinuous','2: Distant:mcont66Other'='Other')
df8$Category<- str_replace_all(df8$Category, rep_str2)
df8 # this is the dataframe for summary results
```

# Adjusted models
```{r}
model1a <- multinom(stage2 ~ In_Medicaid_Flag + race_cat2 + age_cat + Yost_ACS_2006_2010_quintile , data=seer0to19_stage)
summary(model1a)
nrow(fitted(model1a)) # 21,507 is number of obs

b <- c(t(coef(model1a)))
var <- mlogit.clust(model1a, seer0to19_stage, "state")
se <- sqrt(diag(var))

lowCI <- exp(b - 1.96*se)
highCI <-exp(b + 1.96*se)
df9 <- as.data.frame(exp(b))
df10<-as.data.frame(lowCI)
df11<-as.data.frame(highCI)
df12 <- cbind(df9, df10, df11) # this is the result dataset

# relabel columns and drop intercept rows
names(df12)[1:3]<-c("OR", "cluster.lowCI", "cluster.highCI")
df12

df12<-df12[-c(1,3:11, 13:20),]
df12 <- tibble::rownames_to_column(df12, "Category") # Apply rownames_to_column
df12

# Replace multiple strings at a time
rep_str = c('1: Regional:In_Medicaid_FlagEnrolled'='Regional', '2: Distant:In_Medicaid_FlagEnrolled'='Distant')
df12$Stage<- str_replace_all(df12$Category, rep_str)

df12$Category<- "Enrolled"
df12 # this is the dataframe for summary results

# adjusted model
model2a <- multinom(stage2 ~mcont66 + race_cat2 + age_cat + Yost_ACS_2006_2010_quintile, data=seer0to19_stage)
summary(model2a)
nrow(fitted(model2a))

b <- c(t(coef(model2a)))
var <- mlogit.clust(model2a, seer0to19_stage, "state")
se <- sqrt(diag(var))

lowCI <- exp(b - 1.96*se)
highCI <-exp(b + 1.96*se)
df13 <- as.data.frame(exp(b))
df14<-as.data.frame(lowCI)
df15<-as.data.frame(highCI)
df16 <- cbind(df13, df14, df15) # this is the result dataset

# relabel columns and drop intercept rows
names(df16)[1:3]<-c("OR", "cluster.lowCI", "cluster.highCI")
df16

df16<-df16[-c(1,5:13, 17:24),]
df16 <- tibble::rownames_to_column(df16, "Category") # Apply rownames_to_column
df16

# Replace multiple strings at a time
rep_str = c('1: Regional:mcont66Continuous'='Regional','1: Regional:mcont66Discontinuous'='Regional','1: Regional:mcont66Other'='Regional', '2: Distant:mcont66Continuous'='Distant','2: Distant:mcont66Discontinuous'='Distant','2: Distant:mcont66Other'='Distant')
df16$Stage<- str_replace_all(df16$Category, rep_str)

rep_str2 = c('1: Regional:mcont66Continuous'='Continuous','1: Regional:mcont66Discontinuous'='Discontinuous','1: Regional:mcont66Other'='Other', '2: Distant:mcont66Continuous'='Continuous','2: Distant:mcont66Discontinuous'='Discontinuous','2: Distant:mcont66Other'='Other')
df16$Category<- str_replace_all(df16$Category, rep_str2)
df16 # this is the dataframe for summary results
```
# Table 2 with clustered standard errors used for confidence interval calculation
```{r}
Table2 <- rbind(df12, df16)
```

# Relevel to determine if there is a difference between continuous and discontinuous
```{r}
seer0to19_stage <- seer0to19_stage%>%
  mutate(mcont66.2 = case_when(mcont66 == "Continuous" ~ 0,
                               mcont66 == "Discontinuous" ~ 1,
                               mcont66 == "Other" ~ 2,
                               mcont66 == "Not enrolled" ~ 3))
seer0to19_stage$mcont66.2 <- factor(seer0to19_stage$mcont66.2, levels = c(0:3), labels = c("Continuous", "Discontinuous", "Other", "Not enrolled"))

# adjusted model
model2a <- multinom(stage2 ~ mcont66.2 + race_cat2 + age_cat + Yost_ACS_2006_2010_quintile, data=seer0to19_stage)
summary(model2a)
nrow(fitted(model2a))

b <- c(t(coef(model2a)))
var <- mlogit.clust(model2a, seer0to19_stage, "state")
se <- sqrt(diag(var))
z <- b/se
p <- (1 - pnorm(abs(z), 0, 1))*2 # two-tailed z test, reported in manuscript
p

lowCI <- exp(b - 1.96*se)
highCI <-exp(b + 1.96*se)
df17 <- as.data.frame(exp(b))
df18<-as.data.frame(lowCI)
df19<-as.data.frame(highCI)
df20 <- cbind(df17, df18, df19) # this is the result dataset

# relabel columns and drop intercept rows
names(df20)[1:3]<-c("OR", "cluster.lowCI", "cluster.highCI")
df20

df20<-df20[-c(1,5:13, 17:24),]
df20 <- tibble::rownames_to_column(df20, "Category") # Apply rownames_to_column
df20  
```

# SUPPLEMENTARY TABLE 1a by cancer type with clustered standard errors for confidence interval calculation (In_Medicaid_Flag)
```{r}
limited <- seer0to19_stage %>%
 filter(can_type %in% c("I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases","II. Lymphomas and reticuloendothelial neoplasms", "III. CNS and miscellaneous intracranial and intraspinal neoplasms", "IV. Neuroblastoma and other peripheral nervous cell tumors", "V. Retinoblastoma", "VI. Renal tumors", "VII. Hepatic tumors", "VIII. Malignant bone tumors", "IX. Soft tissue and other extraosseous sarcomas", "X. Germ cell tumors, trophoblastic tumors, and neoplasms of gonads", "XI. Other malignant epithelial neoplasms and malignant melanomas"))

ct <- unique(limited$can_type)

# In_Medicaid_Flag
df1a <- data.frame(matrix(ncol=6, nrow=0)) # create the table
columns<-c("Category", "OR","cluster.lowCI", "cluster.highCI", "CancerType") # label the columns with names
colnames(df1a)<-columns

for (i in ct) {
  dfa<- limited %>% 
    filter(can_type == i)
  x<-multinom(stage2 ~ In_Medicaid_Flag + race_cat2 + age_cat + Yost_ACS_2006_2010_quintile, data=dfa)

b <- c(t(coef(x)))
var <- mlogit.clust(x, dfa, "state")
se <- sqrt(diag(var))

lowCI <- exp(b - 1.96*se)
highCI <-exp(b + 1.96*se)
b <- as.data.frame(exp(b))
lCI<-as.data.frame(lowCI)
hCI<-as.data.frame(highCI)
df2a <- cbind(b, lCI, hCI) # this is the result dataset
# relabel columns and dropt intercept rows

names(df2a)[1:3]<-c("OR", "cluster.lowCI", "cluster.highCI")
df2a<-df2a[-c(1,3:11, 13:20),]
df2a <- tibble::rownames_to_column(df2a, "Category") # Apply rownames_to_column
df2a  
df2a$CancerType<-i # add cancer type
df1a <-rbind(df2a, df1a)
}

SuppTable1a<- df1a[order(df1a$CancerType),] # In_Medicaid_Flag

# Mcont66
df3a <- data.frame(matrix(ncol=6, nrow=0)) #create the table
columns<-c("Category", "OR","cluster.lowCI", "cluster.highCI", "CancerType") #label the columns with names
colnames(df3a)<-columns

for (i in ct) {
  dfa<- limited %>% 
    filter(can_type == i)
  x<-multinom(stage2 ~ mcont66 + race_cat2 + age_cat + Yost_ACS_2006_2010_quintile, data=dfa)

b <- c(t(coef(x)))
var <- mlogit.clust(x, dfa, "state")
se <- sqrt(diag(var))

lowCI <- exp(b - 1.96*se)
highCI <-exp(b + 1.96*se)
b <- as.data.frame(exp(b))
lCI<-as.data.frame(lowCI)
hCI<-as.data.frame(highCI)
df4a <- cbind(b, lCI, hCI) # this is the result dataset
# relabel columns and dropt intercept rows

names(df4a)[1:3]<-c("OR", "cluster.lowCI", "cluster.highCI")
df4a <-df4a[-c(1,5:13, 17:24),]
df4a <- tibble::rownames_to_column(df4a, "Category") # Apply rownames_to_column
df4a  
df4a$CancerType<-i # add cancer type
df3a <-rbind(df4a, df3a)
}

SuppTable1b<- df3a[order(df3a$CancerType),] # mcont66
```

# SUPPLEMENTAEY TABLE 1:
```{r}
SuppTable1 <- rbind(SuppTable1a, SuppTable1b)

# Replace multiple strings at a time
rep_str = c('1: Regional:mcont66Continuous'='Regional','1: Regional:mcont66Discontinuous'='Regional','1: Regional:mcont66Other'='Regional', '2: Distant:mcont66Continuous'='Distant','2: Distant:mcont66Discontinuous'='Distant','2: Distant:mcont66Other'='Distant', '1: Regional:In_Medicaid_FlagEnrolled'='Regional','2: Distant:In_Medicaid_FlagEnrolled'='Distant')
SuppTable1$Stage<- str_replace_all(SuppTable1$Category, rep_str)

rep_str2 = c('1: Regional:mcont66Continuous'='Continuous','1: Regional:mcont66Discontinuous'='Discontinuous','1: Regional:mcont66Other'='Other', '2: Distant:mcont66Continuous'='Continuous','2: Distant:mcont66Discontinuous'='Discontinuous','2: Distant:mcont66Other'='Other', '1: Regional:In_Medicaid_FlagEnrolled'='Enrolled','2: Distant:In_Medicaid_FlagEnrolled'='Enrolled')
SuppTable1$Category<- str_replace_all(SuppTable1$Category, rep_str2)

SuppTable1
```

# Add table 2 to SuppTable1 (with clustered SEs) to plot on one plot
```{r}
Table2$CancerType<-"Overall"

SuppTable1 <- rbind(Table2, SuppTable1)
```

# FIGURE 2: Stage analysis by cancer type
```{r}
# graph ORs and 95% CIs
SuppTable1a <- SuppTable1 %>%
  filter(CancerType != "V. Retinoblastoma")

SuppTable1a <- SuppTable1a %>%
  mutate(Abbrev = case_when(CancerType == "Overall" ~ "Overall",
                            CancerType =="II. Lymphomas and reticuloendothelial neoplasms" ~ "LYMPH",
                            CancerType == "III. CNS and miscellaneous intracranial and intraspinal neoplasms" ~ "CNS",
                            CancerType == "IV. Neuroblastoma and other peripheral nervous cell tumors" ~ "NB",
                            CancerType == "V. Retinoblastoma" ~ "RB",
                            CancerType == "VI. Renal tumors" ~ "RT",
                            CancerType == "VII. Hepatic tumors" ~ "HT",
                            CancerType == "VIII. Malignant bone tumors" ~ "MBT",
                            CancerType == "IX. Soft tissue and other extraosseous sarcomas" ~ "STS", 
                            CancerType == "X. Germ cell tumors, trophoblastic tumors, and neoplasms of gonads" ~ "GCT",
                            CancerType == "XI. Other malignant epithelial neoplasms and malignant melanomas" ~ "OTHEP",
                            CancerType == "XII. Other and unspecified malignant neoplasms" ~ "OTHUN"))

SuppTable1a$Abbrev <- factor(SuppTable1a$Abbrev, levels = c("Overall", "LYMPH", "CNS","NB", "RB",
                                                  "RT", "HT", "MBT", "STS", "GCT",
                                                  "OTHEP", "OTHUN"))

SuppTable1a$Stage <- factor(SuppTable1a$Stage, levels = c("Regional", "Distant"))

SuppTable1a$Abbrev<- fct_rev(SuppTable1a$Abbrev)

Plot1 <- SuppTable1a %>%
  filter(Category == "Enrolled", Stage == "Regional") %>%
  ggplot() +
  geom_point(aes(y=OR, x = Abbrev, group = Category, color = Category)) + 
  geom_hline(yintercept =1, linetype=2) +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=OR, x=Abbrev, group = Category, color = Category, ymax = cluster.highCI, ymin = cluster.lowCI), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL, title = "A", x = "Cancer type") +
  ylim(0,2) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) +
  scale_color_manual(breaks = c("Enrolled"), values=c("black"))

Plot1

Plot2 <- SuppTable1a %>%
  filter(Category == "Enrolled", Stage == "Distant") %>%
  ggplot() +
  geom_point(aes(y=OR, x = Abbrev, group = Category, color = Category)) + 
  geom_hline(yintercept =1, linetype=2) +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=OR, x=Abbrev, group = Category, color = Category, ymax = cluster.highCI, ymin = cluster.lowCI), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL, title = "B", x = "Cancer type") +
  ylim(0,3.5) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) +
  scale_color_manual(breaks = c("Enrolled"), values=c("black"))

Plot2

```

# FIGURE 2:  Stage analysis by cancer type mcont66
```{r}
# graph ORs and 95% CIs
SuppTable1a$Category <- factor(SuppTable1a$Category, levels = c("Enrolled", "Discontinuous", "Other", "Continuous"))

Plot3 <- SuppTable1a %>%
  filter(Category != "Enrolled", Stage == "Regional") %>%
  ggplot() +
  geom_point(aes(y=OR, x = Abbrev, group = Category, color = Category), na.rm=TRUE, position=position_dodge(width=0.9)) + 
  geom_hline(yintercept =1, linetype=2) +
  geom_vline(xintercept =c(1.5, 2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5), linetype=2, color = "lightgray") +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(position = position_dodge(width=0.9), aes(y=OR, x=Abbrev, group = Category, color = Category,  ymax = cluster.highCI, ymin = cluster.lowCI),  width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
 # facet_grid(cols = vars(y.level)) +
  labs(color = NULL, title = "C", x = "Cancer type") +
  ylim(0,4.0) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) +
  scale_color_manual(breaks = c("Discontinuous", "Continuous", "Other"),
                        values=c("red", "cyan3", "blue"))

Plot3

Plot4 <- SuppTable1a %>%
  filter(Category != "Enrolled", Stage == "Distant") %>%
  ggplot() +
  geom_point(aes(y=OR, x = Abbrev, group = Category, color = Category), na.rm=TRUE, position=position_dodge(width=0.9)) + 
  geom_hline(yintercept =1, linetype=2) +
  geom_vline(xintercept =c(1.5, 2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5), linetype=2, color = "lightgray") +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(position = position_dodge(width=0.9), aes(y=OR, x=Abbrev, group = Category, color = Category,  max = cluster.highCI, ymin = cluster.lowCI),  width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
 # facet_grid(cols = vars(y.level)) +
  labs(color = NULL, title = "D", x = "Cancer type") +
  ylim(0,5.5) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) +
  scale_color_manual(breaks = c("Discontinuous", "Continuous", "Other"),
                        values=c("red", "cyan3", "blue"))

Plot4
```

```{r}
pdf("Figure 2.pdf", width=8.5, height=11, paper="USr")
Plot1 + Plot2 + Plot3 + Plot4 + plot_layout(ncol=2, nrow=2, guides='collect') &
  theme(legend.position='bottom')
dev.off()
```

# FIGURE 3 survival plots (# In_Medicaid_flag and mcont66) 
```{r}
# In_Medicaid_Flag
plot1 <- ggsurvplot(fit = survfit(Surv(fumo, can_dead) ~ In_Medicaid_Flag, data = seer0to19), 
    xlab = "Months from diagnosis", 
    ylab = "Survival probability", 
    fun="pct",  
    title = "A",
    font.title = c(12, "bold"),
    legend.title = "",
    legend = "bottom",
    font.legend=c(12, "bold"),
    legend.labs = c("Not enrolled", "Enrolled"),
    palette = c("red", "cyan3", "blue", "black"),
    font.x = c(12, "bold"),
    font.y = c(12, "bold"),
    xlim=c(0, 60),
    break.x.by = 12,
    ylim=c(50, 100),
    censor.size=0.5,
    font.tickslab = c(12, "bold"),
    pval=TRUE,
    pval.size = 5,
    pval.coord = c(2,55),
    conf.int = TRUE,
    hline = 90)

# plot1$plot + geom_hline(yintercept = 90)
plot1

# mcont66
 plot2 <- ggsurvplot(fit = survfit(Surv(fumo, can_dead) ~ mcont66, data = seer0to19 ), 
    xlab = "Months from diagnosis", 
    ylab = "Survival probability", 
    fun="pct", 
    title = "B",
    font.title = c(12, "bold"),
    legend.title = "",
    legend = "bottom",
    font.legend=c(12, "bold"),
    legend.labs = c("Not enrolled", "Continuous", "Discontinuous", "Other"),
    palette = c("red", "cyan3", "blue", "black", "gray"),
    font.x = c(12, "bold"),
    font.y = c(12, "bold"),
    xlim=c(0, 60),
    break.x.by = 12,
    ylim=c(50, 100),
    censor.size=0.5,
    font.tickslab = c(12, "bold"),
    pval=TRUE,
    pval.size = 5,
    pval.coord = c(2,55),
    conf.int = TRUE)

plot2
pdf("Figure 3.pdf", width=8.5, height=11, paper="USr")
plot1$plot + plot2$plot + plot_layout(ncol=1, nrow=2, guides='keep') &
  theme(legend.position='bottom')

dev.off()
```
# SUPPLEMENTARY TABLE 2: Estimate survival time 0 to 19 at 60 mo
```{r}
# Overall
summary(survfit(Surv(fumo, can_dead) ~ 1, data = seer0to19), times=60) # estimates 60 month survival time

# In_Medicaid_Flag
SuppTable2a <- summary(survfit(Surv(fumo, can_dead) ~ In_Medicaid_Flag, data = seer0to19), times=60) # estimates 60 month survival time

SuppTable2a <- as.data.frame(cbind(SuppTable2a$strata, SuppTable2a$time, round(SuppTable2a$surv*100, 2), round(SuppTable2a$lower*100, 2), round(SuppTable2a$upper*100,2))) # makes dataframe

SuppTable2a <- setNames(SuppTable2a, c("Category", "Time (months)", "Survival Percentage", "Lower CI", "Upper CI")) # gives column names
SuppTable2a <- SuppTable2a %>%
  mutate(Category = case_when(Category ==1 ~ "Not enrolled",
                              Category ==2 ~ "Enrolled"))

#mcont66
SuppTable2b <- summary(survfit(Surv(fumo, can_dead) ~ mcont66, data = seer0to19), times = 60)
SuppTable2b <- as.data.frame(cbind(SuppTable2b$strata, SuppTable2b$time, round(SuppTable2b$surv*100, 2), round(SuppTable2b$lower*100, 2), round(SuppTable2b$upper*100,2)))

SuppTable2b <- setNames(SuppTable2b, c("Category", "Time (months)", "Survival Percentage", "Lower CI", "Upper CI"))
SuppTable2b <- SuppTable2b %>%
  mutate(Category = case_when(Category ==1 ~ "Not enrolled",
                              Category ==2 ~ "Continuous",
                              Category == 3 ~ "Discontinuous",
                              Category == 4 ~ "Other"))

# Make Supplementary table 2
SuppTable2 <- rbind(SuppTable2a, SuppTable2b)
SuppTable2 <- unique(SuppTable2)
```

# SUPPLEMENTARY TABLE 3: Populate a table with survival by age group at 60 mo
```{r}
age <- unique(seer0to19$age_cat)

# In_Medicaid_Flag
results <- data.frame(matrix(ncol=4, nrow=0)) #create the table
columns<-c("Medicaid Category", "Survival Percentage",  "LowerCI", "UpperCI") # label the columns with names
colnames(results)<-columns

for (i in age) {
  df<- seer0to19 %>% 
    filter(age_cat == i)
  x<-summary(survfit(Surv(fumo, can_dead) ~ In_Medicaid_Flag, df), times=60)
  results2 <- data.frame(matrix(ncol=4, nrow=0)) # create the table
  columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI") #label the columns with names
  colnames(results2)<-columns
  results2[1:2,]<-c(x$strata, round(x$surv,3)*100, round(x$lower,3)*100 , round(x$upper,3)*100)
  results2$Group<- i
  results <-rbind(results, results2)
}

# sort by cancertype and add labels to Medicaid categories
SuppTable3a<- results[order(results$Group),]
SuppTable3a<- SuppTable3a %>%
  rename(Category = Medicaid_Category) %>%
  mutate(Category = case_when(Category == 1 ~ "Not enrolled",
                              Category == 2 ~ "Enrolled"))

# mcont66
# calculate survival at 60 mo
results <- data.frame(matrix(ncol=4, nrow=0)) #create the table
columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI") # label the columns with names
colnames(results)<-columns

for (i in age) {
  df<- seer0to19 %>% 
    filter(age_cat == i)
  x<-summary(survfit(Surv(fumo, can_dead) ~ mcont66, df), times=60)
  results2 <- data.frame(matrix(ncol=4, nrow=0)) # create the table
  columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI") # label the columns with names
  colnames(results2)<-columns
  results2[1:4,]<-c(x$strata, round(x$surv,3)*100, round(x$lower,3)*100 , round(x$upper,3)*100)
  results2$Group<- i
  results <-rbind(results, results2)
}

# sort by group and add labels to Medicaid categories
SuppTable3b <- results[order(results$Group),]
SuppTable3b <- SuppTable3b%>%
  rename(Category = Medicaid_Category) %>%
  mutate(Category = case_when(Category == 1 ~ "Not enrolled",
                              Category == 2 ~ "Continuous",
                              Category == 3 ~ "Discontinuous",
                              Category == 4 ~ "Other"))

SuppTable3a$Group <- factor(SuppTable3a$Group, levels = c("0-4", "5-9", "10-14", "15-19"))
SuppTable3a <- SuppTable3a[order(SuppTable3a$Group),]

SuppTable3b$Group <- factor(SuppTable3b$Group, levels = c("0-4", "5-9", "10-14", "15-19"))
SuppTable3b <- SuppTable3b[order(SuppTable3b$Group),]

SuppTable3 <-rbind(SuppTable3a, SuppTable3b)
SuppTable3 <- unique(SuppTable3) # remove non-unique rows
```

# SUPPLEMENTARY TABLE 4: Populate a table with survival by race at 60 mo
```{r}
race <- unique(seer0to19$race_ethnicity)
limited <- seer0to19 %>%
 filter(race_ethnicity %in% c("Non-Hispanic White","Non-Hispanic Black",  "Non-Hispanic API", "Hispanic", "Non-Hispanic Other"))

# In_Medicaid_Flag
# calculate survival at 60 mo
results <- data.frame(matrix(ncol=4, nrow=0)) # create the table
columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI") #label the columns with names
colnames(results)<-columns

for (i in race) {
  df<- seer0to19 %>% 
    filter(race_ethnicity == i)
  x<-summary(survfit(Surv(fumo, can_dead) ~ In_Medicaid_Flag, df), times=60)
  results2 <- data.frame(matrix(ncol=4, nrow=0)) # create the table
  columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI") # label the columns with names
  colnames(results2)<-columns
  results2[1:2,]<-c(x$strata, round(x$surv,3)*100, round(x$lower,3)*100 , round(x$upper,3)*100)
  results2$Group<- i
  results <-rbind(results, results2)
}

# sort by cancertype and add labels to Medicaid categories
SuppTable4a <- results[order(results$Group),]
SuppTable4a <- SuppTable4a %>%
  rename(Category = Medicaid_Category) %>%
  mutate(Category = case_when(Category == 1 ~ "Not enrolled",
                              Category == 2 ~ "Enrolled"))

# mcont66
# calculate survival at 60 mo
results <- data.frame(matrix(ncol=5, nrow=0)) #create the table
columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI") #label the columns with names
colnames(results)<-columns

for (i in race) {
  df<- limited %>% 
    filter(race_ethnicity == i)
  x<-summary(survfit(Surv(fumo, can_dead) ~ mcont66, df), times=60)
  results2 <- data.frame(matrix(ncol=4, nrow=0)) # create the table
  columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI") # label the columns with names
  colnames(results2)<-columns
  results2[1:4,]<-c(x$strata, round(x$surv,3)*100, round(x$lower,3)*100 , round(x$upper,3)*100)
  results2$Group<- i
  results <-rbind(results, results2)
}

# sort by group and add labels to Medicaid categories
SuppTable4b <- results[order(results$Group),]
SuppTable4b <- SuppTable4b %>%
  rename(Category = Medicaid_Category) %>%
  mutate(Category = case_when(Category == 1 ~ "Not enrolled",
                              Category == 2 ~ "Continuous",
                              Category == 3 ~ "Discontinuous",
                              Category == 4 ~ "Other"))

SuppTable4a$Group <- factor(SuppTable4a$Group, levels = c("Non-Hispanic White", "Non-Hispanic Black", "Non-Hispanic API", "Non-Hispanic Other", "Hispanic"))
SuppTable4a <- SuppTable4a[order(SuppTable4a$Group),]

SuppTable4b$Group <- factor(SuppTable4b$Group, levels = c("Non-Hispanic White", "Non-Hispanic Black", "Non-Hispanic API", "Non-Hispanic Other", "Hispanic"))
SuppTable4b <- SuppTable4b[order(SuppTable4b$Group),]

SuppTable4 <- rbind(SuppTable4a, SuppTable4b)
SuppTable4 <- unique(SuppTable4)
```

# SUPPLEMENTARY TABLE 5: Populate a table with survival by tumor type at 12, 24, 36, 48, and 60 mo
```{r}
limited <- seer0to19 %>%
 filter(can_type %in% c("I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases","II. Lymphomas and reticuloendothelial neoplasms", "III. CNS and miscellaneous intracranial and intraspinal neoplasms", "IV. Neuroblastoma and other peripheral nervous cell tumors", "V. Retinoblastoma", "VI. Renal tumors", "VII. Hepatic tumors", "VIII. Malignant bone tumors", "IX. Soft tissue and other extraosseous sarcomas", "X. Germ cell tumors, trophoblastic tumors, and neoplasms of gonads", "XI. Other malignant epithelial neoplasms and malignant melanomas"))
ct<-unique(limited$can_type)

# In_Medicaid_Flag
# calculate survival at 60 mo
results <- data.frame(matrix(ncol=4, nrow=0)) # create the table
columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI") #label the columns with names
colnames(results)<-columns

for (i in ct) {
  df<- limited %>% 
    filter(can_type == i)
  x<-summary(survfit(Surv(fumo, can_dead) ~ In_Medicaid_Flag, df), times=60)
  results2 <- data.frame(matrix(ncol=4, nrow=0)) # create the table
  columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI") # label the columns with names
  colnames(results2)<-columns
  results2[1:2,]<-c(x$strata, round(x$surv,3)*100, round(x$lower,3)*100 , round(x$upper,3)*100)
  results2$Group<- i
  results <-rbind(results, results2)
}

# sort by cancertype and add labels to Medicaid categories
SuppTable5a <- results[order(results$Group),]
SuppTable5a <- SuppTable5a %>%
  rename(Category = Medicaid_Category) %>%
  mutate(Category = case_when(Category == 1 ~ "Not enrolled",
                              Category == 2 ~ "Enrolled"))

# mcont66
# calculate survival at 60 mo
results <- data.frame(matrix(ncol=4, nrow=0)) # create the table
columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI") # label the columns with names
colnames(results)<-columns

for (i in ct) {
  df<- limited %>% 
    filter(can_type == i)
  x<-summary(survfit(Surv(fumo, can_dead) ~ mcont66, df), times=60)
  results2 <- data.frame(matrix(ncol=4, nrow=0)) # create the table
  columns<-c("Medicaid_Category", "Survival Percentage",  "LowerCI", "UpperCI") #label the columns with names
  colnames(results2)<-columns
  results2[1:4,]<-c(x$strata, round(x$surv,3)*100, round(x$lower,3)*100 , round(x$upper,3)*100)
  results2$Group<- i
  results <-rbind(results, results2)
}
# sort by group and add labels to Medicaid categories
SuppTable5b<- results[order(results$Group),]
SuppTable5b<- SuppTable5b %>%
  rename(Category = Medicaid_Category) %>%
  mutate(Category = case_when(Category == 1 ~ "Not enrolled",
                                       Category == 2 ~ "Continuous",
                                       Category == 3 ~ "Discontinuous",
                                       Category == 4 ~ "Other"))

SuppTable5a$Group <- factor(SuppTable5a$Group, levels = c("I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases","II. Lymphomas and reticuloendothelial neoplasms", "III. CNS and miscellaneous intracranial and intraspinal neoplasms", "IV. Neuroblastoma and other peripheral nervous cell tumors", "V. Retinoblastoma", "VI. Renal tumors", "VII. Hepatic tumors", "VIII. Malignant bone tumors", "IX. Soft tissue and other extraosseous sarcomas", "X. Germ cell tumors, trophoblastic tumors, and neoplasms of gonads", "XI. Other malignant epithelial neoplasms and malignant melanomas"))

SuppTable5a <- SuppTable5a[order(SuppTable5a$Group),]

SuppTable5b$Group <- factor(SuppTable5b$Group, levels = c("I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases","II. Lymphomas and reticuloendothelial neoplasms", "III. CNS and miscellaneous intracranial and intraspinal neoplasms", "IV. Neuroblastoma and other peripheral nervous cell tumors", "V. Retinoblastoma", "VI. Renal tumors", "VII. Hepatic tumors", "VIII. Malignant bone tumors", "IX. Soft tissue and other extraosseous sarcomas", "X. Germ cell tumors, trophoblastic tumors, and neoplasms of gonads", "XI. Other malignant epithelial neoplasms and malignant melanomas"))
SuppTable5b <- SuppTable5b[order(SuppTable5b$Group),]

SuppTable5 <- rbind(SuppTable5a, SuppTable5b)
SuppTable5 <- unique(SuppTable5)
```

# TABLE 3: Cox proportional hazards models (0 to 19) 
```{r}
# unadjusted model 0 TO 19
res.cox1 <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + cluster(state), data = seer0to19) # Univariable model
res.cox1
summary(res.cox1)

# adjusted models 0 to 19
res.cox2 <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19) # Model adjusted for race age and a measure of poverty
res.cox2
summary(res.cox2)

# adjust for stage
res.cox2a <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + stage + cluster(state), data = seer0to19) # Model adjusted for race age and a measure of poverty and stage
res.cox2a
summary(res.cox2a)

# mcont66 unadjusted model 0 TO 19
res.cox3 <- coxph(Surv(fumo, can_dead) ~ mcont66 + cluster(state), data = seer0to19) # Univariable model
res.cox3
summary(res.cox3)

# mcont66 adjusted models 0 to 19
res.cox4 <- coxph(Surv(fumo, can_dead) ~ mcont66 + race_ethnicity +age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19) # Model adjusted for race age and a measure of poverty
res.cox4
summary(res.cox4)

# adjust for stage
res.cox4b <- coxph(Surv(fumo, can_dead) ~ mcont66 + race_ethnicity +age_cat + Yost_ACS_2006_2010_quintile + stage + cluster(state), data = seer0to19) # Model adjusted for race age and a measure of poverty
res.cox4b
summary(res.cox4b)

a<-as.data.frame(exp(cbind(HR = coef(res.cox1), confint(res.cox1))))
   a<-a[c(1),] %>%
    rownames_to_column() %>%
    mutate(model="1. Unadjusted") %>%
    mutate(age_group = "0 to 19 years")

b<-as.data.frame(exp(cbind(HR = coef(res.cox2), confint(res.cox2))))
   b<-b[c(1),] %>%
    rownames_to_column() %>%
    mutate(model="2. Adjusted for race/ethnicity, age, Yost SES indicator") %>%
    mutate(age_group = "0 to 19 years")

c<-as.data.frame(exp(cbind(HR = coef(res.cox3), confint(res.cox3))))
   c<-c[c(1,2,3),] %>%
    rownames_to_column() %>%
    mutate(model="3. Unadjusted") %>%
    mutate(age_group = "0 to 19 years")

d<-as.data.frame(exp(cbind(HR = coef(res.cox4), confint(res.cox4))))
   d<-d[c(1,2,3),] %>%
    rownames_to_column() %>%
    mutate(model="4. Adjusted for race/ethnicity, age, Yost SES indicator") %>%
    mutate(age_group = "0 to 19 years")
Table3 <- rbind(a,b,c,d)  
```

# run a model with continuous as reference
```{r}
seer0to19 <- seer0to19%>%
  mutate(mcont66.2 = case_when(mcont66 == "Continuous" ~ 0,
                               mcont66 == "Discontinuous" ~ 1,
                               mcont66 == "Other" ~ 2,
                               mcont66 == "Not enrolled" ~ 3))

seer0to19$mcont66.2 <- factor(seer0to19$mcont66.2, levels = c(0:3), labels = c("Continuous", "Discontinuous", "Other", "Not enrolled"))

# mcont66.2 adjusted models 0 to 19
res.cox4.2 <- coxph(Surv(fumo, can_dead) ~ mcont66.2 + race_ethnicity +age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19) # Model adjusted for race age and a measure of poverty
res.cox4.2
summary(res.cox4.2)
```

# Test of the ph assumption for adjusted models
```{r}
# adjusted models 0 to 19
Medflag <- cox.zph(res.cox2, terms = FALSE)
Medflag

# mcont66 adjusted models 0 to 19
Med4cat <- cox.zph(res.cox4, terms = FALSE)
Med4cat
# violated for both

plot(Medflag[1], resid =FALSE)
abline(h=0, col=2)

Medflag.surv <- survfit(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity +age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19) # get survival probabilities

plot(Medflag.surv[1], col=c("black", "red"), main="Medicaid_flag log log plot", fun="cloglog", xlab="Time", ylab="log(-log(S(t))") #  ph assumption violated

plot(Med4cat, resid = FALSE)

ggcoxdiagnostics(res.cox2, type = "schoenfeld")

ggcoxdiagnostics(res.cox4, type = "schoenfeld")  
```

# Run models in 12 month increments
```{r}
# Split a survival data set at specified times to form a counting process format
# For those with survival time of 0, add 0.5 mo because otherwise splitting doesn't work and it is more accurate

seer0to19 <- seer0to19 %>%
  mutate(fumo_half = if_else(fumo == 0, 0.5, fumo)) # need to add 0.5 or function doesn't work

seer0to19$patient_id <- as.numeric(rownames(seer0to19))

seer0to19.cp.format <- survSplit(data  = seer0to19,
                            cut   = c(12, 24, 36),       # cut at time 12,24,36
                            end   = "fumo_half",    # original survival time
                            event = "can_dead",   # event indicator
                            start = "start")    # will be created. zero by default

# Recoding
seer0to19.cp.format <- within(seer0to19.cp.format, {
    # Create new survival object
    SurvObj <- Surv(start, fumo_half, can_dead) 
    # Create interval indicator
    interval <- factor(start, levels = c(0,12, 24, 36), labels = c("First","Second", "Third", "Fourth"))
})

# Reordering
seer0to19.cp.format <- seer0to19.cp.format[with(seer0to19.cp.format, order(patient_id, fumo_half)),]  # sort by patient_id and then by fumo
summary(seer0to19$fumo_half)

```

# Supplementary Table 6a: Run models based on intervals for Medicaid_flag
```{r}
age <- unique(seer0to19.cp.format$age_cat)

# testing here
m <-coxph(SurvObj ~ In_Medicaid_Flag + In_Medicaid_Flag*strata(interval) + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + cluster(state), seer0to19.cp.format)

m
library(multcomp)
X<- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1)

t <- glht(m, linfct = X) 
summary(t) # here are the results

CI <- confint(t)
CI <- as.data.frame(exp(CI$confint))
# end testing

W <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1) 
X <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,1,0,0), 1)
Y <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,1,0), 1)
Z <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,1), 1)

x <- coxph(SurvObj ~ In_Medicaid_Flag + In_Medicaid_Flag*strata(interval) + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19.cp.format) 
  t0<- glht(x, linfct = W) 
  df1<-confint(t0)
  df1 <- as.data.frame(exp(df1$confint))

  df1$interval = "0-<12" # enrolled first interval

  t12<- glht(x, linfct = X) 
  df2<-confint(t12)
  df2 <- as.data.frame(exp(df2$confint))
  df2$interval = "12-<24" # enrolled second interval

  t24<- glht(x, linfct = Y) 
  df3<-confint(t24)
  df3 <- as.data.frame(exp(df3$confint))
  df3$interval = "24-<36" # enrolled third interval

  t36 <- glht(x, linfct = Z) 
  df4 <-confint(t36)
  df4 <- as.data.frame(exp(df4$confint))
  df4$interval = "36+" # enrolled fourth interval

  results <-rbind(df1, df2, df3, df4)  # this puts the results from each loop round in the results data file

  results$Group <- "Overall"
  results$category <- "Enrolled"
SuppTable6a <- results

# it seems the risk is highest later on in follow-up but this begs the question about which cancers are driving this?
```

# Graph hazard ratios and confidence intervals across time for enrolled variable
```{r}
p<- results %>%
  ggplot() +
  geom_point(aes(y=Estimate, x= interval), na.rm=TRUE, position=position_dodge(width=0.9), size =4) + 
  geom_hline(yintercept =1, linetype=2) +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=Estimate, x=interval, ymax = lwr, ymin = upr), na.rm=TRUE, position=position_dodge(width=0.9), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() 
p
```
# SUPPLEMENTARY TABLE 6b: Which cancer is driving this? 
```{r}
# In_Medicaid_Flag
results <- data.frame(matrix(ncol=5, nrow=0)) # create empty table for results that will be populated by for loop
columns<-c("estimate", "lwr", "upr", "age", "interval") # label the columns with names
colnames(results) <- columns

cantype <- c("I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases", "III. CNS and miscellaneous intracranial and intraspinal neoplasms")
colnames(seer0to19.cp.format)

for (i in cantype) {
  df <- seer0to19.cp.format %>% filter(can_type == i)

  x <- coxph(SurvObj ~ In_Medicaid_Flag + In_Medicaid_Flag*strata(interval) + age_cat + race_ethnicity + Yost_ACS_2006_2010_quintile + cluster(state), data = df) 
  t0<- glht(x, linfct = W) 
  df1<-confint(t0)
  df1 <- as.data.frame(exp(df1$confint))
  df1$Group <- i
  df1$interval = "0-<12" # enrolled first interval

  t12<- glht(x, linfct = X) 
  df2<-confint(t12)
  df2 <- as.data.frame(exp(df2$confint))
  df2$Group <- i
  df2$interval = "12-<24" # enrolled second interval

  t24<- glht(x, linfct = Y) 
  df3<-confint(t24)
  df3 <- as.data.frame(exp(df3$confint))
  df3$Group <- i
  df3$interval = "24-<36" # enrolled third interval

  t36 <- glht(x, linfct = Z) 
  df4 <-confint(t36)
  df4 <- as.data.frame(exp(df4$confint))
  df4$Group <- i
  df4$interval = "36+" # enrolled fourth interval

  results <-rbind(results, df1, df2, df3, df4)  # this puts the results from each loop round in the results data file
}

results$category = "Enrolled"
SuppTable6b <- results
```

# Group by cancer type and plot
```{r}
SuppTable6b %>%
  group_by(Group) %>%
  ggplot() +
  geom_point(aes(y=Estimate, x=interval)) + 
  geom_hline(yintercept =1, linetype=2) +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=Estimate, x=interval, ymax = lwr, ymin = upr), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() + 
  facet_wrap(~Group, nrow=4, ncol=2, scales = "free") +
  labs( x = "Time interval") + 
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=12, face="bold"))
```
# run models based on intervals for mcont66
```{r}
# testing here
m <-coxph(SurvObj ~ mcont66 + mcont66*strata(interval) + age_cat + race_ethnicity +  Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19.cp.format) 
m

X <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1)

t <- glht(m, linfct = X) 
summary(t) # here are the results

CI<-confint(t)
CI <- as.data.frame(exp(CI$confint))
# end testing

A <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1) # discontinuous 1st interval
B <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1) # other 1st interval
C <- matrix(c(0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1) # Not enrolled 1st interval

D <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0), 1) # discontinuous 2nd interval
E <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0), 1) # other 2nd interval
F <- matrix(c(0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0), 1) # Not enrolled 2nd interval

G <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0), 1) # discontinuous 3rd interval
H <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0), 1) # other 3rd interval
I <- matrix(c(0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0), 1) # Not enrolled 3rd interval

J <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0), 1) # discontinuous 3rd interval
K <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0), 1) # other 3rd interval
L <- matrix(c(0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1), 1) # Not enrolled 3rd interval

  x <- coxph(SurvObj ~ mcont66 + mcont66*strata(interval) + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19.cp.format) 
  t0<- glht(x, linfct = A) 
  df1<-confint(t0)
  df1 <- as.data.frame(exp(df1$confint))
  df1$Group <- "Overall"
  df1$interval = "0-<12"
  df1$category ="continuous" # first interval

  t0<- glht(x, linfct = B) 
  df2<-confint(t0)
  df2 <- as.data.frame(exp(df2$confint))
  df2$Group <- "Overall"
  df2$interval = "0-<12"
  df2$category ="discontinuous" # first interval

  t0<- glht(x, linfct = C) 
  df3<-confint(t0)
  df3 <- as.data.frame(exp(df3$confint))
  df3$Group <- "Overall"
  df3$interval = "0-<12"
  df3$category ="other" # first interval

  t12<- glht(x, linfct = D) 
  df4<-confint(t12)
  df4 <- as.data.frame(exp(df4$confint))
  df4$Group <- "Overall"
  df4$interval = "12-<24"
  df4$category ="continuous" # second interval

  t12<- glht(x, linfct = E) 
  df5<-confint(t12)
  df5 <- as.data.frame(exp(df5$confint))
  df5$Group <- "Overall"
  df5$interval = "12-<24"
  df5$category ="discontinuous" # second interval

  t12<- glht(x, linfct = F) 
  df6<-confint(t12)
  df6 <- as.data.frame(exp(df6$confint))
  df6$Group <- "Overall"
  df6$interval = "12-<24"
  df6$category ="other" # second interval

  t24<- glht(x, linfct = G) 
  df7<-confint(t24)
  df7 <- as.data.frame(exp(df7$confint))
  df7$Group <- "Overall"
  df7$interval = "24-<36"
  df7$category ="continuous" # third interval

  t24<- glht(x, linfct = H) 
  df8<-confint(t24)
  df8<- as.data.frame(exp(df8$confint))
  df8$Group <- "Overall"
  df8$interval = "24-<36"
  df8$category ="discontinuous" # third interval

  t24<- glht(x, linfct = I) 
  df9<-confint(t24)
  df9 <- as.data.frame(exp(df9$confint))
  df9$Group <- "Overall"
  df9$interval = "24-<36"
  df9$category ="other" # third interval

  t36<- glht(x, linfct = J) 
  df10<-confint(t36)
  df10 <- as.data.frame(exp(df10$confint))
  df10$Group <- "Overall"
  df10$interval = "36+"
  df10$category ="continuous" # fourth interval

  t36<- glht(x, linfct = K) 
  df11<-confint(t36)
  df11<- as.data.frame(exp(df11$confint))
  df11$Group <- "Overall"
  df11$interval = "36+"
  df11$category ="discontinuous" # fourth interval

  t36<- glht(x, linfct = L) 
  df12<-confint(t36)
  df12 <- as.data.frame(exp(df12$confint))
  df12$Group <- "Overall"
  df12$interval = "36+"
  df12$category ="other" # fourth interval

  results <-rbind(df1, df2, df3, df4, df5, df6, df7, df8, df9, df10, df11, df12)  # this puts the results from each loop round in the results data file

SuppTable6c <- results
```

# SUPPLEMETARY TABLE 6c: Graph hazard ratios and confidence intervals across time for mcont66 variable
```{r}
# graph HRs and 95% CIs
SuppTable6c %>%
  ggplot() +
  geom_point(aes(y=Estimate, x=interval, group = category, color = category),  position=position_dodge(width=0.9)) + 
  geom_hline(yintercept =1, linetype=2) +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=Estimate, x=interval, ymax = lwr, ymin = upr,  group = category, color = category), position=position_dodge(width=0.9), width = 0.1) +
  geom_vline(xintercept =c(1.5, 2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5), linetype=2, color = "lightgray") +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(y = "HR", x = "Interval") + 
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold"))
```

# SUPPLEMENTARY TABLE 6: HRs by time interval
```{r}
SuppTable6 <- rbind(SuppTable6a, SuppTable6b, SuppTable6c)

```

# Effect modification by sex
```{r}
emsex1 <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + sex + race_ethnicity +age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19) # Model adjusted for race age and a measure of poverty
summary(emsex1)

emsex2 <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + sex + race_ethnicity +age_cat + Yost_ACS_2006_2010_quintile + In_Medicaid_Flag*sex + cluster(state), data = seer0to19) # Model adjusted for race age and a measure of poverty
summary(emsex2)

lrtest(emsex1, emsex2) # not significant
```

# SUPPLEMENTARY TABLE 7a: Test effect modification by age
```{r}
# age_cat*Medicaid flag
res.cox2em <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + In_Medicaid_Flag*age_cat + cluster(state), data = seer0to19) # Model adjusted for race age and a measure of poverty
summary(res.cox2em)

lrtest(res.cox2em, res.cox2)

# significant effect modification. Run models by age group.

cox <-function(agegroup, var){
  x<<-seer0to19 %>%
    filter(age_cat == agegroup)

model<-coxph(Surv(fumo, can_dead) ~ var + race_ethnicity + Yost_ACS_2006_2010_quintile + cluster(state), data = x) # Model adjusted for race  and a measure of poverty

a<-as.data.frame(exp(cbind(HR = coef(model), confint(model))))
   a<-a[c(1),] %>%
    rownames_to_column() %>%
    mutate(model="Adjusted for race_ethnicty and Yost SES indicator") %>%
    mutate(Group = agegroup) 
}

cox2 <-function(agegroup, var){
  x<<-seer0to19 %>%
    filter(age_cat == agegroup)

model<-coxph(Surv(fumo, can_dead) ~ var + race_ethnicity + Yost_ACS_2006_2010_quintile + cluster(state), data = x) # Model adjusted for race  and a measure of poverty

a<-as.data.frame(exp(cbind(HR = coef(model), confint(model))))
   a<-a[c(1,2,3),] %>%
    rownames_to_column() %>%
    mutate(model="Adjusted for race_ethnicty and Yost SES indicator") %>%
    mutate(Group = agegroup) 
}

a<-cox("0-4", x$In_Medicaid_Flag)
b<-cox("5-9", x$In_Medicaid_Flag)
c<-cox("10-14", x$In_Medicaid_Flag)
d<-cox("15-19", x$In_Medicaid_Flag)

e<-cox2("0-4", x$mcont66)
f<-cox2("5-9", x$mcont66)
g<-cox2("10-14", x$mcont66)
h<-cox2("15-19",  x$mcont66)

SuppTable7a <- rbind(a,b,c,d,e,f,g,h)
```

# SUPPLEMENTARY TABLE 7b: Test effect modification by race
```{r}
# race_ethnicity*Medicaid flag
res.cox2emr1 <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to19) # Model adjusted for race age and Yost SES indicator
summary(res.cox2emr1)

res.cox2emr <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + In_Medicaid_Flag*race_ethnicity + cluster(state), data = seer0to19) # Model adjusted for race age and Yost SES indicator
summary(res.cox2emr)

lrtest(res.cox2emr, res.cox2emr1)

# Significant effect modification. Run models by race_ethnicity.
cox <-function(race_e, var){
  x<<-seer0to19 %>%
    filter(race_ethnicity == race_e)

model<-coxph(Surv(fumo, can_dead) ~ var  + age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = x) # Model adjusted for race  and a Yost SES indicator

a<-as.data.frame(exp(cbind(HR = coef(model), confint(model))))
   a<-a[c(1),] %>%
    rownames_to_column() %>%
    mutate(model="Adjusted for age cqtegory and Yost SES indicator") %>%
    mutate(Group = race_e)

}

cox2 <-function(race_e, var){
  x<<-seer0to19 %>%
    filter(race_ethnicity == race_e)

model<-coxph(Surv(fumo, can_dead) ~ var   + age_cat + Yost_ACS_2006_2010_quintile + cluster(state), data = x) # Model adjusted for race  and a measure of poverty

a<-as.data.frame(exp(cbind(HR = coef(model), confint(model))))
   a<-a[c(1,2,3),] %>%
    rownames_to_column() %>%
    mutate(model="Adjusted for race_ethnicty and Yost SES indicator") %>%
    mutate(Group = race_e) 
}

a<-cox("Non-Hispanic White", x$In_Medicaid_Flag) 
b<-cox("Non-Hispanic Black",  x$In_Medicaid_Flag)
c<-cox("Non-Hispanic API", x$In_Medicaid_Flag)
d<-cox("Non-Hispanic Other",  x$In_Medicaid_Flag)
e<-cox("Hispanic", x$In_Medicaid_Flag)

f<-cox2("Non-Hispanic White",  x$mcont66)
g<-cox2("Non-Hispanic Black",x$mcont66)
h<-cox2("Non-Hispanic API", x$mcont66)
i<-cox2("Non-Hispanic Other",  x$mcont66)
j<-cox2("Hispanic", x$mcont66)

SuppTable7b <- rbind(a,b,c,d,e,f,g,h,i,j)

```

# SUPPLEMENTARY TABLE 7c: Test effect modification by cancer type
```{r}
# can_type adjusted
res.cox5 <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + can_type + cluster(state), data = seer0to19) # Model adjusted for race age and Yost SES index
summary(res.cox5)

# can_type*Medicaid flag
res.cox5em <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity + age_cat + Yost_ACS_2006_2010_quintile + can_type + In_Medicaid_Flag*can_type + cluster(state), data = seer0to19) # Model adjusted for race age and Yost SES index
summary(res.cox5em)
lrtest(res.cox5, res.cox5em)

# Significant effect modification. Run models by cancer type.
cox <-function(can_type_e, var){
  x<<-seer0to19 %>%
    filter(can_type == can_type_e)

model<-coxph(Surv(fumo, can_dead) ~ var  + age_cat + race_ethnicity + Yost_ACS_2006_2010_quintile + cluster(state), data = x) # Model adjusted for race, age category, and a measure of poverty

a<-as.data.frame(exp(cbind(HR = coef(model), confint(model))))
   a<-a[c(1),] %>%
     rownames_to_column() %>%
    mutate(model="Adjusted for age cqtegory, race, and Yost SES indicator") %>%
    mutate(Group = can_type_e)

}

cox2 <-function(can_type_e, var){
  x<<-seer0to19 %>%
    filter(can_type == can_type_e)

model<-coxph(Surv(fumo, can_dead) ~ var + age_cat + race_ethnicity + Yost_ACS_2006_2010_quintile + cluster(state), data = x) # Model adjusted for race, age category, and Yost SES index

a<-as.data.frame(exp(cbind(HR = coef(model), confint(model))))
   a<-a[c(1,2,3),] %>%
    rownames_to_column() %>%
   mutate(model="Adjusted for age category, race_ethnicty, and Yost SES index") %>%
    mutate(Group = can_type_e) 
}

a<-cox("I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases", x$In_Medicaid_Flag) 
b<-cox("II. Lymphomas and reticuloendothelial neoplasms",  x$In_Medicaid_Flag)
c<-cox("III. CNS and miscellaneous intracranial and intraspinal neoplasms", x$In_Medicaid_Flag)
d<-cox("IV. Neuroblastoma and other peripheral nervous cell tumors", x$In_Medicaid_Flag)
# e<-cox("V. Retinoblastoma", x$In_Medicaid_Flag) # convergence issues
f<-cox("VI. Renal tumors",x$In_Medicaid_Flag)
g<-cox("VII. Hepatic tumors", x$In_Medicaid_Flag) 
h<-cox("VIII. Malignant bone tumors",  x$In_Medicaid_Flag)
i<-cox("IX. Soft tissue and other extraosseous sarcomas",  x$In_Medicaid_Flag)
j<-cox("X. Germ cell tumors, trophoblastic tumors, and neoplasms of gonads",  x$In_Medicaid_Flag)
k<-cox("XI. Other malignant epithelial neoplasms and malignant melanomas", x$In_Medicaid_Flag)
# l<-cox("XII. Other and unspecified malignant neoplasms", x$In_Medicaid_Flag) # convergence issues

m<-cox2("I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases", x$mcont66)
n<-cox2("II. Lymphomas and reticuloendothelial neoplasms", x$mcont66) 
o<-cox2("III. CNS and miscellaneous intracranial and intraspinal neoplasms", x$mcont66)
p<-cox2("IV. Neuroblastoma and other peripheral nervous cell tumors", x$mcont66)
# q<-cox2("V. Retinoblastoma", x$mcont66)# convergence issues
r<-cox2("VI. Renal tumors", x$mcont66)
s<-cox2("VII. Hepatic tumors", x$mcont66)
t<-cox2("VIII. Malignant bone tumors", x$mcont66)
u<-cox2("IX. Soft tissue and other extraosseous sarcomas", x$mcont66)
v<-cox2("X. Germ cell tumors, trophoblastic tumors, and neoplasms of gonads", x$mcont66)
w<-cox2("XI. Other malignant epithelial neoplasms and malignant melanomas", x$mcont66)
# y<-cox2("XII. Other and unspecified malignant neoplasms", x$mcont66) # convergence issues

SuppTable7c <- rbind(a,b,c,d,f,h,i,j,k, m,n,o,p,r,s,t,u,v,w)

```

# SUPPLEMENTARY TABLE 7: Effect modifcation table
```{r}
# combine age, race/ethnicity, and cancer type specific results
SuppTable7<- rbind(SuppTable7a, SuppTable7b, SuppTable7c)
table(SuppTable7$rowname)

SuppTable7 <- SuppTable7 %>%
  rename(Category = rowname) %>%
  mutate(Category = case_when(Category == "varContinuous" ~ "Continuous",
            Category == "varDiscontinuous" ~ "Discontinuous",
            Category == "varOther" ~ "Other",
            Category == "varEnrolled" ~ "Enrolled")) %>%
  mutate(Group2 = case_when(Group %in% c("0-4", "5-9", "10-14", "15-19") ~ "Age",
                            Group %in% c("I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases", "II. Lymphomas and reticuloendothelial neoplasms", "III. CNS and miscellaneous intracranial and intraspinal neoplasms", "IV. Neuroblastoma and other peripheral nervous cell tumors", "V. Retinoblastoma", "VI. Renal tumors", "VII. Hepatic tumors", "VIII. Malignant bone tumors", "IX. Soft tissue and other extraosseous sarcomas",  "X. Germ cell tumors, trophoblastic tumors, and neoplasms of gonads", "XI. Other malignant epithelial neoplasms and malignant melanomas", "XII. Other and unspecified malignant neoplasms") ~ "CancerType",
                            Group %in% c("Non-Hispanic White", "Non-Hispanic Black", "Non-Hispanic API", "Non-Hispanic Other", "Hispanic") ~ "Race"))

# need abbreviations for plot fitting

SuppTable7 <- SuppTable7 %>%
  mutate(Abbrev = case_when(Group == "I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases" ~ "LEUK",
                            Group =="II. Lymphomas and reticuloendothelial neoplasms" ~ "LYMPH",
                            Group == "III. CNS and miscellaneous intracranial and intraspinal neoplasms" ~ "CNS",
                            Group == "IV. Neuroblastoma and other peripheral nervous cell tumors" ~ "NB",
                            Group == "V. Retinoblastoma" ~ "RB",
                            Group == "VI. Renal tumors" ~ "RT",
                            Group == "VII. Hepatic tumors" ~ "HT",
                            Group == "VIII. Malignant bone tumors" ~ "MBT",
                            Group == "IX. Soft tissue and other extraosseous sarcomas" ~ "STS", 
                            Group == "X. Germ cell tumors, trophoblastic tumors, and neoplasms of gonads" ~ "GCT",
                            Group == "XI. Other malignant epithelial neoplasms and malignant melanomas" ~ "OTHEP",
                            Group == "XII. Other and unspecified malignant neoplasms" ~ "OTHUN",
                            Group == "Non-Hispanic White" ~ "NHW",
                            Group == "Non-Hispanic Black" ~ "NHB",
                            Group == "Non-Hispanic API" ~ "NHAPI",
                            Group == "Non-Hispanic Other"~ "NHOTH",
                            Group == "Hispanic" ~ "HISP",
                            Group == "0-4" ~ "0 to 4",
                            Group == "5-9" ~ "5 to 9",
                            Group == "10-14" ~ "10 to 14",
                            Group == "15-19" ~ "15 to 19"))

SuppTable7$Abbrev <- factor(SuppTable7$Abbrev, levels = c("LEUK", "LYMPH", "CNS","NB", "RB",
                                                  "RT", "HT", "MBT", "STS", "GCT",
                                                  "OTHEP", "OTHUN", "NHW", "NHB",  "NHAPI",
                                                  "NHOTH", "HISP", "0 to 4", "5 to 9", "10 to 14", "15 to 19"))
SuppTable7$Abbrev <- fct_rev(SuppTable7$Abbrev)
```

# Supplementary FIGURE 1: Effect modification by age, race, cancer type
- Medicaid enrollment variable
```{r}
plot1 <- SuppTable7 %>%
  filter(Category == "Enrolled", Group2 == "Age") %>%
ggplot() +
  geom_point(aes(y=HR, x = Abbrev, group=Category, color = Category)) + 
  geom_hline(yintercept =1, linetype=2) +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=HR, x=Abbrev, group = Category, color = Category, ymax = `2.5 %`, ymin = `97.5 %`), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL, title = " ", x = "") +
  ylim(0.5,3) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) + 
          ggtitle("A") + theme(plot.title = element_text(face="bold")) +
  scale_color_manual(breaks = c("Enrolled"), values=c("black")) +
  theme(legend.title=element_blank())
plot1

plot2 <-SuppTable7 %>%
  filter(Category == "Enrolled", Group2 == "Race") %>% 
  ggplot() +
  geom_point(aes(y=HR, x = Abbrev, group = Category, color = Category)) + 
  geom_hline(yintercept =1, linetype=2) +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=HR, x=Abbrev, group = Category, color = Category, ymax = `2.5 %`, ymin = `97.5 %`), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL, title = " ", x = "") +
  ylim(0.5,5) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) + ggtitle("C") +theme(plot.title = element_text(face="bold")) +
  scale_color_manual(breaks = c("Enrolled"),
                        values=c("black")) +
  theme(legend.title=element_blank())

plot2

plot3 <-SuppTable7 %>%
  filter(Category == "Enrolled", Group2 == "CancerType", Abbrev != "OTHUN") %>%
  ggplot() +
  geom_point(aes(y=HR, x = Abbrev, group = Category, color = Category)) + 
  geom_hline(yintercept =1, linetype=2) +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=HR, x=Abbrev, group = Category, color = Category, ymax = `2.5 %`, ymin = `97.5 %`), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL, title = " ", x = "") +
  ylim(0.5,3.6) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) + ggtitle("E") +theme(plot.title = element_text(face="bold")) +
  scale_color_manual(breaks = c("Enrolled"),
                        values=c("black")) +
  theme(legend.title=element_blank())

plot3
```
# Supplementary FIGURE 1: Effect modification by age, race, cancer type
- mcont66
```{r}
SuppTable7$Category <- factor(SuppTable7$Category, levels = c("Enrolled", "Discontinuous", "Other",  "Continuous"))

plot4 <- SuppTable7 %>%
  filter(Category != "Enrolled", Group2 == "Age") %>%
ggplot() +
  geom_point(aes(y=HR, x = Abbrev, group=Category, color = Category), na.rm=TRUE, position=position_dodge(width = 0.9)) + 
  geom_hline(yintercept =1, linetype=2) +
   geom_vline(xintercept = c(1.5, 2.5,3.5), linetype=2, color = "lightgray") +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(position=position_dodge(width = 0.9), aes(y=HR, x=Abbrev, group=Category, color = Category, ymax = `2.5 %`, ymin = `97.5 %`), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL, title = " ", x = "") +
  ylim(0.5,3) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) + ggtitle("B") +theme(plot.title = element_text(face="bold")) +
  scale_color_manual(breaks = c("Discontinuous", "Continuous", "Other"),
                        values=c("red", "cyan3", "blue")) +
  theme(legend.title=element_blank())

plot4

plot5 <-SuppTable7 %>%
  filter(Category != "Enrolled", Group2 == "Race") %>% 
  ggplot() +
  geom_point(aes(y=HR, x = Abbrev, group=Category, color = Category), na.rm=TRUE, position=position_dodge(width = 0.9)) + 
  geom_hline(yintercept =1, linetype=2) +
  geom_vline(xintercept =c(1.5, 2.5,3.5, 4.5), linetype=2, color = "lightgray") +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(position=position_dodge(width = 0.9), aes(y=HR, x=Abbrev, group=Category, color = Category, ymax = `2.5 %`, ymin = `97.5 %`), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL,title = " ", x = "") +
  ylim(0.5,7) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) + ggtitle("D") +theme(plot.title = element_text(face="bold")) +
  scale_color_manual(breaks = c("Discontinuous", "Continuous", "Other"),
                        values=c("red", "cyan3", "blue")) +
  theme(legend.title=element_blank())

plot5

plot6 <-SuppTable7 %>%
  filter(Category != "Enrolled", Group2 == "CancerType", !Abbrev %in% c("RB", "OTHUN")) %>%
  ggplot() +
  geom_point(aes(y=HR, x = Abbrev, group=Category, color = Category), na.rm=TRUE, position=position_dodge(width = 0.9)) + 
  geom_hline(yintercept =1, linetype=2) +
  geom_vline(xintercept =c(1.5, 2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5), linetype=2, color = "lightgray") +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(position=position_dodge(width = 0.9), aes(y=HR, x=Abbrev, group=Category, color = Category, ymax = `2.5 %`, ymin = `97.5 %`), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL, title = " ", x = "") +
  ylim(0.5,4.5) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) + ggtitle("F") +theme(plot.title = element_text(face="bold"))+
  scale_color_manual(breaks = c("Discontinuous", "Continuous", "Other"),
                        values=c("red", "cyan3", "blue"))+
  theme(legend.title=element_blank())

plot6
```
# Supplementary FIGURE 1: Export
```{r}
pdf("Supplementary Figure 1.pdf", width=12, height=10, paper="USr")
plot1 + plot2 + plot3 + plot4 + plot5 + plot6 + plot_layout(ncol=3, nrow=2, guides='collect') &
  theme(legend.position='bottom')

dev.off()
```

# Clean up tables
```{r}
Table2 <- Table2%>%
  rename(`lower 95% CI` = cluster.lowCI,
         `upper 95% CI` = cluster.highCI) %>%
  mutate_if(is.numeric, round, 2)

Table2$`OR (95% CI)` <-paste0(Table2$OR, " (", Table2$`lower 95% CI`, " to ", Table2$`upper 95% CI`, ")")

Table3 <- Table3 %>% 
  rename(Category = rowname,
         "lower 95% CI" = "2.5 %",
         "upper 95% CI" = "97.5 %") %>%
  mutate(Category = case_when(Category == "In_Medicaid_FlagEnrolled" ~ "Enrolled",
            Category == "mcont66Discontinuous" ~ "Discontinuous",
            Category == "mcont66Other" ~ "Other",
            Category == "mcont66Continuous" ~ "Continuous"))  %>%
  mutate_if(is.numeric, round, 2)

Table3$`HR (95% CI)` <-paste0(Table3$HR, " (", Table3$`lower 95% CI`, " to ", Table3$`upper 95% CI`, ")") 
```

# Supplementary Tables
```{r}
# 1
SuppTable1 <- SuppTable1 %>%
  rename(`lower 95% CI` = cluster.lowCI,
         `upper 95% CI` = cluster.highCI,
          `Cancer Type` = CancerType) %>%
  mutate_if(is.numeric, round, 2)

SuppTable1$`OR (95% CI)` <-paste0(SuppTable1$OR, " (", SuppTable1$`lower 95% CI`, " to ", SuppTable1$`upper 95% CI`, ")") 
SuppTable1 <- SuppTable1 %>%
  dplyr::select(1,5,6,7)

# 2
SuppTable2 <- SuppTable2 %>%
  rename(`lower 95% CI` = `Lower CI`,
         `upper 95% CI` = `Upper CI`)

SuppTable2$`Survival Percentage (95% CI)` <-paste0(SuppTable2$`Survival Percentage`, " (", SuppTable2$`lower 95% CI`, " to ", SuppTable2$`upper 95% CI`, ")")
# drop uneeded columns
SuppTable2 <- SuppTable2 %>%
  dplyr::select(1,5)

# 3
SuppTable3 <- SuppTable3 %>%
  rename(`lower 95% CI` = LowerCI,
         `upper 95% CI` = UpperCI)

SuppTable3$`Survival Percentage (95% CI)` <-paste0(SuppTable3$`Survival Percentage`, " (", SuppTable3$`lower 95% CI`, " to ", SuppTable3$`upper 95% CI`, ")")
SuppTable3 <- SuppTable3 %>%
  dplyr::select(1,5,6)

# 4
SuppTable4 <- SuppTable4 %>%
  rename(`lower 95% CI` = LowerCI,
         `upper 95% CI` = UpperCI)

SuppTable4$`Survival Percentage (95% CI)` <-paste0(SuppTable4$`Survival Percentage`, " (", SuppTable4$`lower 95% CI`, " to ", SuppTable4$`upper 95% CI`, ")")
SuppTable4 <- SuppTable4 %>%
  dplyr::select(1,5,6)

# 5
SuppTable5 <- SuppTable5 %>%
  rename(`lower 95% CI` = LowerCI,
         `upper 95% CI` = UpperCI)

SuppTable5$`Survival Percentage (95% CI)` <-paste0(SuppTable5$`Survival Percentage`, " (", SuppTable5$`lower 95% CI`, " to ", SuppTable5$`upper 95% CI`, ")")

SuppTable5 <- SuppTable5 %>%
  dplyr::select(1,5,6)

# 6
SuppTable6 <- SuppTable6 %>%
  rename(`Time interval` = interval,
         Category = category,
         `lower 95% CI` = lwr,
         `upper 95% CI` = upr) %>%
  mutate(Category = case_when(Category == "Enrolled"~ "Enrolled",
            Category == "discontinuous" ~ "Discontinuous",
            Category == "other" ~ "Other",
            Category == "continuous" ~ "Continuous")) %>%
  mutate_if(is.numeric, round, 2)

SuppTable6$`HR (95% CI)` <-paste0(SuppTable6$Estimate, " (", SuppTable6$`lower 95% CI`, " to ", SuppTable6$`upper 95% CI`, ")") 
SuppTable6 <- SuppTable6 %>%
  dplyr::select(4,5,6,7)

# 7
SuppTable7 <- SuppTable7 %>%
  rename(`lower 95% CI` = `2.5 %`,
         `upper 95% CI` = `97.5 %`) %>%
  dplyr::select(-7, -8)  %>%
  mutate_if(is.numeric, round, 2)

SuppTable7$`HR (95% CI)` <-paste0(SuppTable7$HR, " (", SuppTable7$`lower 95% CI`, " to ", SuppTable7$`upper 95% CI`, ")") 
SuppTable7 <- SuppTable7 %>%
  dplyr::select(1,6,7)
```

# Export data to excel (Tables 2-3)
```{r}
sheets<-list("Table 2 (unadj stagedx enro)" = df4,
             "Table 2 (unadj stagedx mcont66)" = df8,
             "Table 2 (stage dx analysis)" = Table2,
             "Table 3 (overall CPH)" = Table3)
write.xlsx(sheets, file="SEER0to19analysis_tables.xlsx", keepNA=TRUE, overwrite=TRUE)
```

# Export supplementary tables to excel
```{r}
sheets <- list("Supp Table 1 (stage cantype)" = SuppTable1,
               "Supp Table 2 (surv overall)" = SuppTable2,
               "Supp Table 3 (surv prob age )" = SuppTable3,
               "Supp Table 4 (surv prob race )" = SuppTable4,
               "Supp Table 5 (surv prob can )" = SuppTable5,
               "Supp Table 6 (ph analysis)" = SuppTable6, 
               "Supp Table 7 (EM)" = SuppTable7)
write.xlsx(sheets, file="Supplementary tables updated.xlsx", showNA=TRUE, overwrite = TRUE)
```

# SOME Sensitivity analyses
# This code validates the function above for clustered SEs using binomial logistic (nnet object is hard to work with for any standard functions that calculate robust or clustered SEs) 
```{r}
# regional vs. in situ/localized
model1_logit<-glm(stage ~ In_Medicaid_Flag, data = seer0to19_stage[which(seer0to19_stage$stage %in% c("In situ/localized", "Regional")),], family = binomial(link = "logit"))
odds.n.ends(model1_logit) # these are CIs without robust SEs

# distant vs. in situ/localized
model1_logit2<-glm(stage ~ In_Medicaid_Flag, data = seer0to19_stage[which(seer0to19_stage$stage %in% c("In situ/localized", "Distant")),], family = binomial(link = "logit"))
odds.n.ends(model1_logit2)  # these are CIs without robust SEs

model1_logit3 <-glm(stage ~ mcont66, data = seer0to19_stage[which(seer0to19_stage$stage %in% c("In situ/localized", "Distant")),], family = binomial(link = "logit"))
odds.n.ends(model1_logit3)  # these are CIs without robust SEs

model1_logit4 <-glm(stage ~ mcont66, data = seer0to19_stage[which(seer0to19_stage$stage %in% c("In situ/localized", "Distant")),], family = binomial(link = "logit"))
odds.n.ends(model1_logit4)  # these are CIs without robust SEs
```

# get results with clustered standard errors (solution 5  here: https://stackoverflow.com/questions/16498849/logistic-regression-with-robust-clustered-standard-errors-in-r)
```{r}
model1_logit_RSE <- miceadds::glm.cluster(data=seer0to19_stage[which(seer0to19_stage$stage %in% c("In situ/localized", "Regional")),], formula= stage ~ In_Medicaid_Flag + race_cat2 + age_cat + Yost_ACS_2006_2010_quintile, cluster="state", family="binomial")
summary(model1_logit_RSE)
exp(0.156116634)
exp(0.156116634 + (1.96*0.02744999)) 
exp(0.156116634 - (1.96*0.02744999))

model2_logit_RSE <- miceadds::glm.cluster(data=seer0to19_stage[which(seer0to19_stage$stage %in% c("In situ/localized", "Distant")),], formula= stage ~ In_Medicaid_Flag + race_cat2 + age_cat + Yost_ACS_2006_2010_quintile, cluster="state", family="binomial")
summary(model2_logit_RSE)

exp(0.354153776)
exp(0.354153776 + (1.96*0.02011365)) 
exp(0.354153776 - (1.96*0.02011365))

# The results are the same.
```